

**10.41**: _Clustering via \(k\)-means as an approximate matrix factorization._ Suppose we run the \(k\)-means algorithm on the \(N\)\(n\)-vectors \(x_{1},\ldots,x_{N}\), to obtain the group representatives \(z_{1},\ldots,z_{k}\). Define the matrices \(X=[\begin{array}{cccc}x_{1}&\cdots&x_{N}\end{array}]\), \(Z=[\begin{array}{cccc}z_{1}&\cdots&z_{N}\end{array}]\). \(X\) has size \(n\times N\) and \(Z\) has size \(n\times k\). We encode the assignment of vectors to groups by the \(k\times N\) clustering matrix \(C\), with \(C_{ij}=1\) if \(x_{j}\) is assigned to group \(i\), and \(C_{ij}=0\) otherwise. Each column of \(C\) is a unit vector; its transpose is a selector matrix. 1. Give an interpretation of the columns of the matrix \(X-ZC\), and the squared norm \(\|X-ZC\|^{2}\). 2. Justify the following statement: The goal of the \(k\)-means algorithm is to find an \(n\times k\) matrix \(Z\), and a \(k\times N\) matrix \(C\), which is the transpose of a selector matrix, so that \(\|X-ZC\|\) is small, _i.e._, \(X\approx ZC\).
**10.42**: A matrix-vector multiplication \(Ax\) of an \(n\times n\) matrix \(A\) and an \(n\)-vector \(x\) takes \(2n^{2}\) flops in general. Formulate a faster method, with complexity linear in \(n\), for matrix-vector multiplication with the matrix \(A=I+ab^{T}\), where \(a\) and \(b\) are given \(n\)-vectors.
**10.43**: A particular computer takes about 0.2 seconds to multiply two \(1500\times 1500\) matrices. About how long would you guess the computer takes to multiply two \(3000\times 3000\) matrices? Give your prediction (_i.e._, the time in seconds), and your (very brief) reasoning.
**10.44**: _Complexity of matrix quadruple product._ (See page 182.) We wish to compute the product \(E=ABCD\), where \(A\) is \(m\times n\), \(B\) is \(n\times p\), \(C\) is \(p\times q\), and \(D\) is \(q\times r\). 1. Find all methods for computing \(E\) using three matrix-matrix multiplications. For example, you can compute \(AB\), \(CD\), and then the product \((AB)(CD)\). Give the total number of flops required for each of these methods. _Hint._ There are four other methods. 2. Which method requires the fewest flops, with dimensions \(m=10\), \(n=1000\), \(p=10\), \(q=1000\), \(r=100\)?