* _Running sum matrix_. The \(n\times n\) matrix \[S=\left[\begin{array}{cccccc}1&0&0&\cdots&0&0\\ 1&1&0&\cdots&0&0\\ &&\ddots&\ddots&&\\ &&&\ddots&\ddots&\\ 1&1&1&\cdots&1&0\\ 1&1&1&\cdots&1&1\end{array}\right]\] (6.6) is called the _running sum matrix_. The \(i\)th entry of the \(n\)-vector \(Sx\) is the sum of the first \(i\) entries of \(x\): \[Sx=\left[\begin{array}{c}x_{1}\\ x_{1}+x_{2}\\ x_{1}+x_{2}+x_{3}\\ \vdots\\ x_{1}+\cdots+x_{n}\end{array}\right].\]

Application examples.
* _Feature matrix and weight vector._ Suppose \(X\) is a feature matrix, where its \(N\) columns \(x_{1},\ldots,x_{N}\) are feature \(n\)-vectors for \(N\) objects or examples. Let the \(n\)-vector \(w\) be a _weight vector_, and let \(s_{i}=x_{i}^{T}w\) be the score associated with object \(i\) using the weight vector \(w\). Then we can write \(s=X^{T}w\), where \(s\) is the \(N\)-vector of scores of the objects.
* _Portfolio return time series._ Suppose that \(R\) is a \(T\times n\) asset return matrix, that gives the returns of \(n\) assets over \(T\) periods. A common trading strategy maintains constant investment weights given by the \(n\)-vector \(w\) over the \(T\) periods. For example, \(w_{4}=0.15\) means that \(15\%\) of the total portfolio value is held in asset 4. (Short positions are denoted by negative entries in \(w\).) Then \(Rw\), which is a \(T\)-vector, is the time series of the portfolio returns over the periods \(1,\ldots,T\). As an example, consider a portfolio of the 4 assets in table 6.1, with weights \(w=(0.4,0.3,-0.2,0.5)\). The product \(Rw=(0.00213,-0.00201,0.00241)\) gives the portfolio returns over the three periods in the example.
* _Polynomial evaluation at multiple points._ Suppose the entries of the \(n\)-vector \(c\) are the coefficients of a polynomial \(p\) of degree \(n-1\) or less: \[p(t)=c_{1}+c_{2}t+\cdots+c_{n-1}t^{n-2}+c_{n}t^{n-1}.\] Let \(t_{1},\ldots,t_{m}\) be \(m\) numbers, and define the \(m\)-vector \(y\) as \(y_{i}=p(t_{i})\). Then we have \(y=Ac\), where \(A\) is the \(m\times n\) matrix \[A=\left[\begin{array}{ccccc}1&t_{1}&\cdots&t_{1}^{n-2}&t_{1}^{n-1}\\ 1&t_{2}&\cdots&t_{2}^{n-2}&t_{2}^{n-1}\\ \vdots&\vdots&&\vdots&\vdots\\ 1&t_{m}&\cdots&t_{m}^{n-2}&t_{m}^{n-1}\end{array}\right].\] (6.7) 