only holds for \(\beta_{1}=\cdots=\beta_{k}=0\). In other words, the only linear combination of the vectors that equals the zero vector is the linear combination with all coefficients zero.

As with linear dependence, we will say "The vectors \(a_{1},\ldots,a_{k}\) are linearly independent" to mean "The list of vectors \(a_{1},\ldots,a_{k}\) is linearly independent". But, like linear dependence, linear independence is an attribute of a collection of vectors, and not individual vectors.

It is generally not easy to determine by casual inspection whether or not a list of vectors is linearly dependent or linearly independent. But we will soon see an algorithm that does this.

#### Examples.

* A list consisting of a single vector is linearly dependent only if the vector is zero. It is linearly independent only if the vector is nonzero.
* Any list of vectors containing the zero vector is linearly dependent.
* A list of two vectors is linearly dependent if and only if one of the vectors is a multiple of the other one. More generally, a list of vectors is linearly dependent if any one of the vectors is a multiple of another one.
* The vectors \[a_{1}=\left[\begin{array}{c}0.2\\ -7.0\\ 8.6\end{array}\right],\qquad a_{2}=\left[\begin{array}{c}-0.1\\ 2.0\\ -1.0\end{array}\right],\qquad a_{3}=\left[\begin{array}{c}0.0\\ -1.0\\ 2.2\end{array}\right]\] are linearly dependent, since \(a_{1}+2a_{2}-3a_{3}=0\). We can express any of these vectors as a linear combination of the other two. For example, we have \(a_{2}=(-1/2)a_{1}+(3/2)a_{3}\).
* The vectors \[a_{1}=\left[\begin{array}{c}1\\ 0\\ 0\end{array}\right],\qquad a_{2}=\left[\begin{array}{c}0\\ -1\\ 1\end{array}\right],\qquad a_{3}=\left[\begin{array}{c}-1\\ 1\\ 1\end{array}\right]\] are linearly independent. To see this, suppose \(\beta_{1}a_{1}+\beta_{2}a_{2}+\beta_{3}a_{3}=0\). This means that \[\beta_{1}-\beta_{3}=0,\qquad-\beta_{2}+\beta_{3}=0,\qquad\beta_{2}+\beta_{3}=0.\] Adding the last two equations we find that \(2\beta_{3}=-0\), so \(\beta_{3}=0\). Using this, the first equation is then \(\beta_{1}=0\), and the second equation is \(\beta_{2}=0\).
* The standard unit \(n\)-vectors \(e_{1},\ldots,e_{n}\) are linearly independent. To see this, suppose that (5.1) holds. We have \[0=\beta_{1}e_{1}+\cdots+\beta_{n}e_{n}=\left[\begin{array}{c}\beta_{1}\\ \vdots\\ \beta_{n}\end{array}\right],\] so we conclude that \(\beta_{1}=\cdots=\beta_{n}=0\).

 