

**39.**: From \(m\) independent measurements \(b_{1},\ldots,b_{m}\) of your pulse rate, weighted by \(w_{1},\ldots,w_{m}\), what is the weighted average that replaces equation (9)? It is the best estimate when the statistical variances are \(\sigma_{i}^{2}\equiv 1/w_{i}^{2}\).
**40.**: If \(W=\left[\begin{smallmatrix}2&0\\ 0&1\end{smallmatrix}\right]\), find the \(W\)-inner product of \(x=(2,3)\) and \(y=(1,1)\), and the \(W\)-length of \(x\). What line of vectors is \(W\)-perpendicular to \(y\)?
**41.**: Find the weighted least-squares solution \(\widehat{x}_{W}\) to \(Ax=b\):

\[A=\begin{bmatrix}1&0\\ 1&1\\ 1&2\end{bmatrix}\qquad b=\begin{bmatrix}0\\ 1\\ 1\end{bmatrix}\qquad W=\begin{bmatrix}2&0&0\\ 0&1&0\\ 0&0&1\end{bmatrix}.\]

Check that the projection \(A\widehat{x}_{W}\) is still perpendicular (in the \(W\)-inner product!) to the error \(b-A\widehat{x}_{W}\).
**42.**:
* Suppose you guess your professor's age, making errors \(e=-2,-1,5\) with probabilities \(\frac{1}{2},\frac{1}{4},\frac{1}{4}\). Check that the expected error \(E(e)\) is zero and find the variance \(E(e^{2})\).
* If the professor guesses too (or tries to remember), making errors \(-1\), \(0\), \(1\) with probabilities \(\frac{1}{8},\frac{6}{8},\frac{1}{8}\), what weights \(w_{1}\) and \(w_{2}\) give the reliability of your guess and the professor's guess?

### Orthogonal Bases and Gram-Schmidt

In an orthogonal basis, every vector is perpendicular to every other vector. The coordinate axes are mutually orthogonal. That is just about optimal, and the one possible improvement is easy: Divide each vector by its length, to make it a _unit vector_. That changes an _orthogonal_ basis into an _orthonormal_ basis of \(q\)'s:

**3P**: The vectors \(q_{1},\ldots,q_{n}\) are _orthonormal_ if

\[q_{i}^{\mathrm{T}}q_{j}=\begin{cases}0&\text{whenever $i\neq j$},\qquad \text{giving the orthogonality;}\\ 1&\text{whenever $i=j$},\qquad\text{giving the normalization.}\end{cases}\]

_A matrix with orthonormal columns will be called Q._

The most important example is the _standard basis_. For the \(x\)-\(y\) plane, the best-known axes \(e_{1}=(1,0)\) and \(e_{2}=(0,1)\) are not only perpendicular but horizontal and vertical. \(Q\) is the 2 by 2 identity matrix. In \(n\) dimensions the standard basis \(e_{1},\ldots,e_{n}\) again consists