Show that \(A\) cannot be invertible. The third row of \(A^{-1}\), multiplying \(A\), should give the third row \([0\ \ 0\ \ 1\ \ 0]\) of \(A^{-1}A=I\). Why is this impossible?
**10.**: Find the inverses (in any legal way) of

\[A_{1}=\begin{bmatrix}0&0&0&1\\ 0&0&2&0\\ 0&3&0&0\\ 4&0&0&0\end{bmatrix},\qquad A_{2}=\begin{bmatrix}1&0&0&0\\ -\frac{1}{2}&1&0&0\\ 0&-\frac{2}{3}&1&0\\ 0&0&-\frac{3}{4}&1\end{bmatrix},\qquad A_{3}=\begin{bmatrix}a&b&0&0\\ c&d&0&0\\ 0&0&a&b\\ 0&0&c&d\end{bmatrix}.\]
**11.**: Give examples of \(A\) and \(B\) such that

1. \(A+B\) is not invertible although \(A\) and \(B\) are invertible.
2. \(A+B\) is invertible although \(A\) and \(B\) are not invertible.
3. all of \(A\), \(B\), and \(A+B\) are invertible.
4. In the last case use \(A^{-1}(A+B)B^{-1}=B^{-1}+A^{-1}\) to show that \(C=B^{-1}+A^{-1}\) is also invertible--and find a formula for \(C^{-1}\).
**12.**: If \(A\) is invertible, which properties of \(A\) remain true for \(A^{-1}\)?

(a) \(A\) is triangular. (b) \(A\) is symmetric. (c) \(A\) is tridiagonal. (d) All entries are whole numbers. (e) All entries are fractions (including numbers like \(\frac{3}{1}\)).
**13.**: If \(A=[\begin{smallmatrix}3\\ 1\end{smallmatrix}]\) and \(B=[\begin{smallmatrix}2\\ 2\end{smallmatrix}]\), compute \(A^{\rm T}B\), \(B^{\rm T}A\), \(AB^{\rm T}\), and \(BA^{\rm T}\).
**14.**: If \(B\) is square, show that \(A=B+B^{\rm T}\) is always symmetric and \(K=B-B^{\rm T}\) is always _skew-symmetric_--which means that \(K^{\rm T}=-K\). Find these matrices \(A\) and \(K\) when \(B=[\begin{smallmatrix}1&3\\ 1&1\end{smallmatrix}]\), and write \(B\) as the sum of a symmetric matrix and a skew-symmetric matrix.
**15.**:
1. How many entries can be chosen independently in a symmetric matrix of order \(n\)?
2. How many entries can be chosen independently in a skew-symmetric matrix (\(K^{\rm T}=-K\)) of order \(n\)? The diagonal of \(K\) is zero!
**16.**:
1. If \(A=LDU\), with 1s on the diagonals of \(L\) and \(U\), what is the corresponding factorization of \(A^{\rm T}\)? Note that \(A\) and \(A^{\rm T}\) (square matrices with no row exchanges) share the same pivots.
2. What triangular systems will give the solution to \(A^{\rm T}y=b\)?
**17.**: If \(A=L_{1}D_{1}U_{1}\) and \(A=L_{2}D_{2}U_{2}\), prove that \(L_{1}=L_{2}\), \(D_{1}=D_{2}\), and \(U_{1}=U_{2}\). If \(A\) is invertible, the factorization is unique.

1. Derive the equation \(L_{1}^{-1}L_{2}D_{2}=D_{1}U_{1}U_{2}^{-1}\), and explain why one side is lower triangular and the other side is upper triangular.
2. Compare the main diagonals and then compare the off-diagonals.

 