

**31.**: _Tridiagonal matrices_ have zero entries except on the main diagonal and the two adjacent diagonals. Factor these into \(A=LU\) and \(A=LDV\):

\[A=\begin{bmatrix}1&1&0\\ 1&2&1\\ 0&1&2\end{bmatrix}\quad\text{and}\quad A=\begin{bmatrix}a&a&0\\ a&a+b&b\\ 0&b&b+c\end{bmatrix}.\]
**32.**: Solve the triangular system \(Lc=b\) to find \(c\). Then solve \(Ux=c\) to find \(x\):

\[L=\begin{bmatrix}1&0\\ 4&1\end{bmatrix}\quad\text{and}\quad U=\begin{bmatrix}2&4\\ 0&1\end{bmatrix}\quad\text{and}\quad b=\begin{bmatrix}2\\ 11\end{bmatrix}.\]

For safety find \(A=LU\) and solve \(Ax=b\) as usual. Circle \(c\) when you see it.
**33.**: Solve \(Lc=b\) to find \(c\). Then solve \(Ux=c\) to find \(x\). What was \(A\)?

\[L=\begin{bmatrix}1&0&0\\ 1&1&0\\ 1&1&1\end{bmatrix}\quad\text{and}\quad U=\begin{bmatrix}1&1&1\\ 0&1&1\\ 0&0&1\end{bmatrix}\quad\text{and}\quad b=\begin{bmatrix}4\\ 5\\ 6\end{bmatrix}.\]
**34.**: If \(A\) and \(B\) have nonzeros in the positions marked by \(x\), which zeros are still zero in their factors \(L\) and \(U\)?

\[A=\begin{bmatrix}x&x&x&x\\ x&x&x&0\\ 0&x&x&x\\ 0&0&x&x\end{bmatrix}\quad\text{and}\quad B=\begin{bmatrix}x&x&x&0\\ x&x&0&x\\ x&0&x&x\\ 0&x&x&x\end{bmatrix}.\]
**35.**: (Important) If \(A\) has pivots 2, 7, 6 with no row exchanges, what are the pivots for the upper left 2 by 2 submatrix \(B\) (without row 3 and column 3)? Explain why.
**36.**: Starting from a 3 by 3 matrix \(A\) with pivots 2, 7, 6, add a fourth row and column to produce \(M\). What are the first three pivots for \(M\), and why? What fourth row and column are sure to produce 9 as the fourth pivot?
**37.**: Use chol(pascal(5)) to find the triangular factors of MATLAB's pascal(5). Row exchanges in [L, U] = lu(pascal(5)) spoil Pascal's pattern!
**38.**: (Review) For which numbers \(c\) is \(A=LU\) impossible--with three pivots?

\[A=\begin{bmatrix}1&2&0\\ 3&c&1\\ 0&1&1\end{bmatrix}.\]
**39.**: Estimate the time difference for each new right-hand side \(b\) when \(n=800\). Create A = rand(800) and b = rand(800,1) and B = rand(800,9). Compare the times from tic; A\(\backslash\)b; toc and tic; A\(\backslash\)B; toc (which solves for 9 right sides).
**Problems 40-48 are about permutation matrices.**: