

**15.**: For four linear equations in two unknowns \(x\) and \(y\), the row picture shows four . The column picture is in .-dimensional space. The equations have no solution unless the vector on the right-hand side is a combination of .
**16.**: Find a point with \(z=2\) on the intersection line of the planes \(x+y+3z=6\) and \(x-y+z=4\). Find the point with \(z=0\) and a third point halfway between.
**17.**: The first of these equations plus the second equals the third:

\[\begin{array}{ccccccc}x&+&y&+&z&=&2\\ x&+&2y&+&z&=&3\\ 2x&+&3y&+&2z&=&5.\end{array}\]

The first two planes meet along a line. The third plane contains that line, because if \(x\), \(y\), \(z\) satisfy the first two equations then they also . The equations have infinitely many solutions (the whole line \(\mathbf{L}\)). Find three solutions.
**18.**: Move the third plane in Problem 17 to a parallel plane \(2x+3y+2z=9\). Now the three equations have no solution--_why not_? The first two planes meet along the line \(\mathbf{L}\), but the third plane doesn't that line.
**19.**: In Problem 17 the columns are \((1,1,2)\) and \((1,2,3)\) and \((1,1,2)\). This is a "singular case" because the third column is . Find two combinations of the columns that give \(b=(2,3,5)\). This is only possible for \(b=(4,6,c)\) if .
**20.**: Normally 4 "planes" in four-dimensional space meet at a . Normally 4 column vectors in four-dimensional space can combine to produce \(b\). What combination of \((1,0,0,0)\), \((1,1,0,0)\), \((1,1,1,0)\), \((1,1,1,1)\) produces \(b=(3,3,3,2)\)? What 4 equations for \(x\), \(y\), \(z\), \(t\) are you solving?
**21.**: When equation 1 is added to equation 2, which of these are changed: the planes in the row picture, the column picture, the coefficient matrix, the solution?
**22.**: If \((a,b)\) is a multiple of \((c,d)\) with \(abcd\neq 0\), _show that \((a,c)\) is a multiple of \((b,d)\)_. This is surprisingly important: call it a challenge question. You could use numbers first to see how \(a\), \(b\), \(c\), and \(d\) are related. The question will lead to:

If \(A=\left[\begin{smallmatrix}a&b\\ c&d\end{smallmatrix}\right]\) has dependent rows then it has dependent columns.
**23.**: In these equations, the third column (multiplying \(w\)) is the _same_ as the right side \(b\). The column form of the equations _immediately_ gives what solution for \((u,v,w)\)?

\[\begin{array}{ccccccc}6u&+&7v&+&8w&=&8\\ 4u&+&5v&+&9w&=&9\\ 2u&-&2v&+&7w&=&7.\end{array}\]