

**5.** Which of these matrices cannot be diagonalized?

\[A_{1}=\begin{bmatrix}2&-2\\ 2&-2\end{bmatrix}\qquad A_{2}=\begin{bmatrix}2&0\\ 2&-2\end{bmatrix}\qquad A_{3}=\begin{bmatrix}2&0\\ 2&2\end{bmatrix}.\]

**6.**:
1. If \(A^{2}=I\), what are the possible eigenvalues of \(A\)? 2. If this \(A\) is 2 by 2, and not \(I\) or \(-I\), find its trace and determinant. 3. If the first row is \((3,-1)\), what is the second row?
**7.**: If \(A=\begin{bmatrix}4&3\\ 1&2\end{bmatrix}\), find \(A^{100}\) by diagonalizing \(A\).
**8.**: Suppose \(A=w^{\rm T}\) is a column times a row (a rank-1 matrix).

1. By multiplying \(A\) times \(u\), show that \(u\) is an eigenvector. What is \(\lambda\)? 2. What are the other eigenvalues of \(A\) (and why)? 3. Compute \({\rm trace}(A)\) from the sum on the diagonal and the sum of \(\lambda\)'s.
**9.**: Show by direct calculation that \(AB\) and \(BA\) have the same trace when

\[A=\begin{bmatrix}a&b\\ c&d\end{bmatrix}\qquad\text{and}\qquad B=\begin{bmatrix}q&r\\ s&t\end{bmatrix}.\]

 Deduce that \(AB-BA=I\) is impossible (except in infinite dimensions).
**10.**: Suppose \(A\) has eigenvalues 1, 2, 4. What is the trace of \(A^{2}\)? What is the determinant of \((A^{-1})^{\rm T}\)?
**11.**: If the eigenvalues of \(A\) are 1, 1, 2, which of the following are certain to be true? Give a reason if true or a counterexample if false:

1. \(A\) is invertible.
2. \(A\) is diagonalizable.
3. \(A\) is not diagonalizable.
**12.**: Suppose the only eigenvectors of \(A\) are multiples of \(x=(1,0,0)\). True or false:

1. \(A\) is not invertible.
2. \(A\) has a repeated eigenvalue.
3. \(A\) is not diagonalizable.
**13.**: Diagonalize the matrix \(A=\begin{bmatrix}5&4\\ 4&5\end{bmatrix}\) and find one of its square roots--a matrix such that \(R^{2}=A\). How many square roots will there be?
**14.**: Suppose the eigenvector matrix \(S\) has \(S^{\rm T}=S^{-1}\). Show that \(A=S\Lambda S^{-1}\) is symmetric and has orthogonal eigenvectors.

**Problems 15-24 are about the eigenvalue and eigenvector matrices.**