

**34.**: Following the method of Problem 33, reduce \(A\) to echelon form and look at zero rows. The \(b\) column tells which combinations you have taken of the rows:

\[\text{(a)}\quad\begin{bmatrix}1&2&b_{1}\\ 3&4&b_{2}\\ 4&6&b_{3}\end{bmatrix}.\qquad\qquad\qquad\text{(b)}\quad\begin{bmatrix}1&2&b_{1 }\\ 2&3&b_{2}\\ 2&4&b_{3}\\ 2&5&b_{4}\end{bmatrix}.\]

From the \(b\) column after elimination, read off \(m-r\) basis vectors in the left nullspace of \(A\) (combinations of rows that give zero).
**35.**: Suppose \(A\) is the sum of two matrices of rank one: \(A=uv^{\text{T}}+wz^{\text{T}}\).

1. Which vectors span the column space of \(A\)?
2. Which vectors span the row space of \(A\)?
3. The rank is less than 2 if or if .
4. Compute \(A\) and its rank if \(u=z=(1,0,0)\) and \(v=w=(0,0,1)\).
**36.**: Without multiplying matrices, find bases for the row and column spaces of \(A\):

\[A=\begin{bmatrix}1&2\\ 4&5\\ 2&7\end{bmatrix}\begin{bmatrix}3&0&3\\ 1&1&2\end{bmatrix}.\]

How do you know from these shapes that \(A\) is not invertible?
**37.**: True or false (with a reason or a counterexample)?

1. \(A\) and \(A^{\text{T}}\) have the same number of pivots.
2. \(A\) and \(A^{\text{T}}\) have the same left nullspace.
3. If the row space equals the column space then \(A^{\text{T}}=A\).
4. If \(A^{\text{T}}=-A\) then the row space of \(A\) equals the column space.
**38.**: If \(AB=0\), the columns of \(B\) are in the nullspace of \(A\). If those vectors are in \(\mathbf{R}^{n}\), prove that \(\operatorname{rank}(A)+\operatorname{rank}(B)\leq n\).
**39.**: Can tic-tac-toe be completed (5 ones and 4 zeros in \(A\)) so that \(\operatorname{rank}(A)=2\) but neither side passed up a winning move?
**40.**: Construct any 2 by 3 matrix of rank 1. Copy Figure 2.5 and put one vector in each subspace (two in the nullspace). Which vectors are orthogonal?
**41.**: Redraw Figure 2.5 for a 3 by 2 matrix of rank \(r=2\). Which subspace is \(Z\) (zero vector only)? The nullspace part of any vector \(x\) in \(\mathbf{R}^{2}\) is \(x_{n}=\)