Find a 3 by 3 system with these solutions exactly when \(b_{1}+b_{2}=b_{3}\).
**11.**: Write a 2 by 2 system \(Ax=b\) with many solutions \(x_{n}\) but no solution \(x_{p}\). (Therefore the system has no solution.) Which \(b\)'s allow an \(x_{p}\)?
**12.**: Which of these rules give a correct definition of the _rank_ of \(A\)?

1. The number of nonzero rows in \(R\).
2. The number of columns minus the total number of rows.
3. The number of columns minus the number of free columns.
4. The number of 1s in \(R\).
**13.**: Find the reduced row echelon forms \(R\) and the rank of these matrices:

1. The 3 by 4 matrix of all 1s.
2. The 4 by 4 matrix with \(a_{ij}=(-1)^{ij}\).
3. The 3 by 4 matrix with \(a_{ij}=(-1)^{j}\).
**14.**: Find \(R\) for each of these (block) matrices, and the special solutions:

\[A=\begin{bmatrix}0&0&0\\ 0&0&3\\ 2&4&6\end{bmatrix}\qquad B=\begin{bmatrix}A&A\end{bmatrix}\qquad C=\begin{bmatrix} A&A\\ A&0\end{bmatrix}.\]
**15.**: If the \(r\) pivot variables come first, the reduced \(R\) must look like

\[R=\begin{bmatrix}I&F\\ 0&0\end{bmatrix}\qquad\begin{array}{l}I\text{ is $r$ by $r$}\\ \text{$F$ is $r$ by $n-r$}\end{array}\]

What is the nullspace matrix \(N\) containing the special solutions?
**16.**: Suppose all \(r\) pivot variables come _last_. Describe the four blocks in the \(m\) by \(n\) reduced echelon form (the block \(B\) should be \(r\) by \(r\)):

\[R=\begin{bmatrix}A&B\\ C&D\end{bmatrix}.\]

What is the nullspace matrix \(N\) of special solutions? What is its shape?
**17.**: (Silly problem) Describe all 2 by 3 matrices \(A_{1}\) and \(A_{2}\) with row echelon forms \(R_{1}\) and \(R_{2}\), such that \(R_{1}+R_{2}\) is the row echelon form of \(A_{1}+A_{2}\). Is it true that \(R_{1}=A_{1}\) and \(R_{2}=A_{2}\) in this case?
**18.**: If \(A\) has \(r\) pivot columns, then \(A^{\mathrm{T}}\) has \(r\) pivot columns. Give a 3 by 3 example for which the column numbers are different for \(A\) and \(A^{\mathrm{T}}\).

 