1. How do you know that the rows are linearly dependent? 2. Explain why all 120 terms are zero in the big formula for \(\det A\).
**17.** Find two ways to choose nonzeros from four different rows and columns: \[A=\begin{bmatrix}1&0&0&1\\ 0&1&1&1\\ 1&1&0&1\\ 1&0&0&1\end{bmatrix}\qquad B=\begin{bmatrix}1&0&0&2\\ 0&3&4&5\\ 5&4&0&3\\ 2&0&0&1\end{bmatrix}.\quad(B\text{ has the same zeros as }A.)\] Is \(\det A\) equal to \(1+1\) or \(1-1\) or \(-1-1\)? What is \(\det B\)?
**18.** Place the smallest number of zeros in a 4 by 4 matrix that will guarantee \(\det A=0\). Place as many zeros as possible while still allowing \(\det A\neq 0\).
**19.**: 1. If \(a_{11}=a_{22}=a_{33}=0\), how many of the six terms in \(\det A\) will be zero? 2. If \(a_{11}=a_{22}=a_{33}=a_{44}=0\), how many of the 24 products \(a_{1j}a_{2k}a_{3\ell}a_{4m}\) are sure to be zero?
**20.** How many 5 by 5 permutation matrices have \(\det P=+1\)? Those are even permutations. Find one that needs four exchanges to reach the identity matrix.
**21.** If \(\det A\neq 0\), at least one of the \(n!\) terms in the big formula (6) is not zero. Deduce that some ordering of the rows of \(A\) leaves no zeros on the diagonal. (Don't use \(P\) from elimination; that \(PA\) can have zeros on the diagonal.)
**22.** Prove that 4 is the largest determinant for a 3 by 3 matrix of 1s and \(-1\)s.
**23.** How many permutations of \((1,2,3,4)\) are even and what are they? Extra credit: What are all the possible 4 by 4 determinants of \(I+P_{\text{even}}\)?

**Problems 24-33 use cofactors \(C_{ij}=(-1)^{i+j}\det M_{ij}\). Delete row \(i\), column \(j\).**
**24.** Find cofactors and then transpose. Multiply \(C_{A}^{\text{T}}\) and \(C_{B}^{\text{T}}\) by \(A\) and \(B\)!

\[A=\begin{bmatrix}2&1\\ 3&6\end{bmatrix}\qquad B=\begin{bmatrix}1&2&3\\ 4&5&6\\ 7&0&0\end{bmatrix}.\]
**25.** Find the cofactor matrix \(C\) and compare \(AC^{\text{T}}\) with \(A^{-1}\):

\[A=\begin{bmatrix}2&-1&0\\ -1&2&-1\\ 0&-1&2\end{bmatrix}\qquad A^{-1}=\frac{1}{4}\begin{bmatrix}3&2&1\\ 2&4&2\\ 1&2&3\end{bmatrix}.\] 