the plane of columns 2 and 3 . Only two columns are independent.
The vector $b=(2,5,7)$ is in that plane of the columns-it is column 1 plus column 3 -so $(1,0,1)$ is a solution. We can add an multiple of the combination $(3,-1,-2)$ that gives $b=0$. So there is a whole line of solutions-as we know from the row picture.

The truth is that we knew the columns would combine to give zero, because the rows did. That is a fact of mathematics, not of computation-and it remains true in dimension $n$. If the n planes have no point in common, or infinitely many points, then the $n$ columns lie in the same plane.

If the row picture breaks down, so does the column picture. That brings out the difference between Chapter 1 and Chapter 2. This chapter studies the most important problem-the nonsingular case-where there is one solution and it has to be found. Chapter 2 studies the general case, where there may be many solutions or none. In both cases we cannot continue without a decent notation (matrix notation) and a decent algorithm (elimination). After the exercises, we start with elimination.

Problem Set 1.2
1. For the equations $x+y=4,2 x-2 y=4$, draw the row picture (two intersecting lines) and the column picture (combination of two columns equal to the column vector $(4,4)$ on the right side).
2. Solve to find a combination of the columns that equals $b$ :
Triangular system
$$
\begin{aligned}
u-v-w & =b_1 \\
v+w & =b_2 \\
w & =b_3 .
\end{aligned}
$$
3. (Recommended) Describe the intersection of the three planes $u+v+w+z=6$ and $u+w+z=4$ and $u+w=2$ (all in four-dimensional space). Is it a line or a point or an empty set? What is the intersection if the fourth plane $u=-1$ is included? Find a fourth equation that leaves us with no solution.
4. Sketch these three lines and decide if the equations are solvable:
3 by 2 system
$$
\begin{aligned}
x+2 y & =2 \\
x-y & =2 \\
y & =1 .
\end{aligned}
$$

What happens if all right-hand sides are zero? Is there any nonzero choice of righthand sides that allows the three lines to intersect at the same point?
5. Find two points on the line of intersection of the three planes $t=0$ and $z=0$ and $x+y+z+t=1$ in four-dimensional space.
