Which previous problem justifies the middle step?

**7.**: By choosing the correct vector \(b\) in the Schwarz inequality, prove that

\[(a_{1}+\cdots+a_{n})^{2}\leq n(a_{1}^{2}+\cdots+a_{n}^{2}).\]

When does equality hold?
**8.**: The methane molecule CH\({}_{4}\) is arranged as if the carbon atom were at the center of a regular tetrahedron with four hydrogen atoms at the vertices. If vertices are placed at \((0,0,0)\), \((1,1,0)\), \((1,0,1)\), and \((0,1,1)\)--note that all six edges have length \(\sqrt{2}\), so the tetrahedron is regular--what is the cosine of the angle between the rays going from the center \((\frac{1}{2},\frac{1}{2},\frac{1}{2})\) to the vertices? (The bond angle itself is about \(109.5^{\circ}\), an old friend of chemists.)
**9.**: Square the matrix \(P=aa^{\rm T}/a^{\rm T}a\), which projects onto a line, and show that \(P^{2}=P\). (Note the number \(a^{\rm T}a\) in the middle of the matrix \(aa^{\rm T}aa^{\rm T}\)!)
**10.**: Is the projection matrix \(P\) invertible? Why or why not?
**11.**: 1. Find the projection matrix \(P_{1}\) onto the line through \(a=[^{1}_{3}]\) and also the matrix \(P_{2}\) that projects onto the line perpendicular to \(a\). 2. Compute \(P_{1}+P_{2}\) and \(P_{1}P_{2}\) and explain.
**12.**: Find the matrix that projects every point in the plane onto the line \(x+2y=0\).
**13.**: Prove that the _trace_ of \(P=aa^{\rm T}/a^{\rm T}a\)--which is the sum of its diagonal entries--always equals 1.
**14.**: What matrix \(P\) projects every point in \({\bf R}^{3}\) onto the line of intersection of the planes \(x+y+t=0\) and \(x-t=0\)?
**15.**: Show that the length of \(Ax\) equals the length of \(A^{\rm T}x\) if \(AA^{\rm T}=A^{\rm T}A\).
**16.**: Suppose \(P\) is the projection matrix onto the line through \(a\).

1. Why is the inner product of \(x\) with \(Py\) equal to the inner product of \(Px\) with \(y\)?
2. Are the two angles the same? Find their cosines if \(a=(1,1,-1)\), \(x=(2,0,1)\), \(y=(2,1,2)\).
3. Why is the inner product of \(Px\) with \(Py\) again the same? What is the angle between those two?
**Problems 17-26 ask for projections onto lines. Also errors \(e=b-p\) and matrices \(P\).**
**17.**: Project the vector \(b\) onto the line through \(a\). Check that \(e\) is perpendicular to \(a\) 