* How is the determinant of \(A^{\rm H}\) related to the determinant of \(A\)?
* Prove that the determinant of any Hermitian matrix is real.
* How many degrees of freedom are there in a real symmetric matrix, a real diagonal matrix, and a real orthogonal matrix? (The first answer is the sum of the other two, because \(A=Q\Lambda Q^{\rm T}\).)
* Show that 3 by 3 Hermitian matrices \(A\) and also unitary \(U\) have 9 real degrees of freedom (columns of \(U\) can be multiplied by any \(e^{i\theta}\)).
* Write \(P\), \(Q\) and \(R\) in the form \(\lambda_{1}x_{1}x_{1}^{\rm H}+\lambda_{2}x_{2}x_{2}^{\rm H}\) of the spectral theorem: \[P=\begin{bmatrix}\frac{1}{2}&\frac{1}{2}\\ \frac{1}{2}&\frac{1}{2}\end{bmatrix},\qquad Q=\begin{bmatrix}0&1\\ 1&0\end{bmatrix},\qquad R=\begin{bmatrix}3&4\\ 4&-3\end{bmatrix}.\]
* Give a reason if true or a counterexample if false: 1. If \(A\) is Hermitian, then \(A+iI\) is invertible. 2. If \(Q\) is orthogonal. then \(Q+\frac{1}{2}I\) is invertible. 3. If \(A\) is real, then \(A+iI\) is invertible.
* Suppose \(A\) is a symmetric 3 by 3 matrix with eigenvalues 0, 1, 2. 1. What properties can be guaranteed for the corresponding unit eigenvectors \(u\), \(v\), \(w\)? 2. In terms of \(u\), \(v\), \(w\), describe the nullspace, left nullspace, row space and column space of \(A\). 3. Find a vector \(x\) that satisfies \(Ax=v+w\). Is \(x\) unique? 4. Under what conditions on \(b\) does \(Ax=b\) have a solution? 5. If \(u\), \(v\), \(w\) are the columns of \(S\), what are \(S^{-1}\) and \(S^{-1}AS\)?
* In the list below, which classes of matrices contain \(A\) and which contain \(B\)? \[A=\begin{bmatrix}0&1&0&0\\ 0&0&1&0\\ 0&0&0&1\\ 1&0&0&0\end{bmatrix}\qquad\text{ and }\qquad B=\frac{1}{4}\begin{bmatrix}1&1&1&1 \\ 1&1&1&1\\ 1&1&1&1\\ 1&1&1&1\end{bmatrix}.\] _Orthogonal, invertible, projection, permutation, Hermitian, rank-\(1\), diagonalizable, Markov_. Find the eigenvalues of \(A\) and \(B\).
* What is the dimension of the space \(S\) of all \(n\) by \(n\) real symmetric matrices? The spectral theorem says that every symmetric matrix is a combination of \(n\) projection matrices. Since the dimension exceeds \(n\), how is this difference explained?
* Write one significant fact about the eigenvalues of each of the following.

 