2. Those vectors (are)(are not)(might be) linearly independent. 3. Any four of those vectors (are)(are not)(might be) a basis for \({\bf R}^{4}\). 4. If those vectors are the columns of \(A\), then \(Ax=b\) (has) (does not have) (might not have) a solution.
**23.**: The columns of \(A\) are \(n\) vectors from \({\bf R}^{m}\). If they are linearly independent, what is the rank of \(A\)? If they span \({\bf R}^{m}\), what is the rank? If they are a basis for \({\bf R}^{m}\), what then?
**24.**: Find a basis for the plane \(x-2y+3z=0\) in \({\bf R}^{3}\). Then find a basis for the intersection of that plane with the \(xy\)-plane. Then find a basis for all vectors perpendicular to the plane.
**25.**: Suppose the columns of a 5 by 5 matrix \(A\) are a basis for \({\bf R}^{5}\).

1. The equation \(Ax=0\) has only the solution \(x=0\) because .
2. If \(b\) is in \({\bf R}^{5}\) then \(Ax=b\) is solvable because .

Conclusion: \(A\) is invertible. Its rank is 5.
**26.**: Suppose \({\bf S}\) is a five-dimensional subspace of \({\bf R}^{6}\). True or false?

1. Every basis for \({\bf S}\) can be extended to a basis for \({\bf R}^{6}\) by adding one more vector.
2. Every basis for \({\bf R}^{6}\) can be reduced to a basis for \({\bf S}\) by removing one vector.
**27.**: \(U\) comes from \(A\) by subtracting row 1 from row 3: \[A=\begin{bmatrix}1&3&2\\ 0&1&1\\ 1&3&2\end{bmatrix}\quad\text{and}\quad U=\begin{bmatrix}1&3&2\\ 0&1&1\\ 0&0&0\end{bmatrix}.\] Find bases for the two column spaces. Find bases for the two nullspace.
**28.**: True or false (give a good reason)?

1. If the columns of a matrix are dependent, so are the rows.
2. The column space of a 2 by 2 matrix is the same as its row space.
3. The column space of a 2 by 2 matrix has the same dimension as its row space.
4. The columns of a matrix are a basis for the column space.
**29.**: For which numbers \(c\) and \(d\) do these matrices have rank 2?

\[A=\begin{bmatrix}1&2&5&0&5\\ 0&0&c&2&2\\ 0&0&0&d&2\end{bmatrix}\quad\text{and}\quad B=\begin{bmatrix}c&d\\ d&c\end{bmatrix}.\] 