

**3.**: If you could increase the capacity of any one pipe in the network above, which change would produce the largest increase in the maximal flow?
**4.**: Draw a 5-node network with capacity \(|i-j|\) between node \(i\) and node \(j\). Find the largest possible flow from node 1 to node 4.
**5.**: In a graph, the maximum number of paths from \(s\) to \(t\) with no common edges equals the minimum number of edges whose removal disconnects \(s\) from \(t\). Relate this to the max flow-min cut theorem.
**6.**: Find a maximal set of marriages (a complete matching, if possible) for \[A=\begin{bmatrix}0&0&1&0&0\\ 1&1&0&1&1\\ 0&1&1&0&1\\ 0&0&1&1&0\\ 0&0&0&1&0\end{bmatrix}\qquad\text{and}\qquad B=\begin{bmatrix}1&1&0&0&0\\ 0&1&0&1&0\\ 0&0&1&0&1\\ 1&1&1&0&0\\ 1&0&0&0&0\end{bmatrix}.\] Sketch the network for \(B\), with heavier lines on the edges in your matching.
**7.**: For the matrix \(A\) in Problem 6, which rows violate Hall's condition--by having all their 1s in too few columns? Which \(p\) by \(q\) submatrix of zeros has \(p+q>n\)?
**8.**: How many lines (horizontal and vertical) are needed to cover all the 1s in \(A\) in Problem 6? For any matrix, explain why weak duality is true: If \(k\) marriages are possible, then it takes at least \(k\) lines to cover all the 1s.
**9.**: 1. Suppose every row and every column contains exactly two 1s. Prove that a complete matching is possible. (Show that the 1s cannot be covered by less than \(n\) lines)
**2.**: Find an example with two or _more_ is in each row and column, for which a complete matching is impossible.
**10.**: If a 7 by 7 matrix has 15 1s, prove that it allows at least 3 marriages.
**11.**: For _infinite_ sets, a complete matching may be impossible even if Hall's condition is passed. If the first row is all 1s and then every \(a_{i\,i-1}=1\), show that any \(p\) rows have 1s in at least \(p\) columns--and yet there is no complete matching.

