Can you find a time \(T\) at which the solution \(u(T)\) is guaranteed to return to the initial value \(u(0)\)?
* If \(P\) is the matrix that projects \(\mathbf{R}^{n}\) onto a subspace \(\mathbf{S}\), explain why every vector in \(\mathbf{S}\) is an eigenvector, and so is every vector in \(\mathbf{S}^{\perp}\). What are the eigenvai (Note the connection to \(P^{2}=P\), which means that \(\lambda^{2}=\lambda\).)
* Show that every matrix of order \(>1\) is the sum of two singular matrices.
* Show that the matrix differential equation \(dX/dt=AX+XB\) has the solution \(X(t)=e^{At}X(0)e^{Bt}\).
* Prove that the solutions of \(dX/dt=AX-XA\) keep the same eigenvalues for all time.
* If the eigenvalues of \(A\) are \(1\) and \(3\) with eigenvectors \((5,2)\) and \((2,1)\), find the solutions to \(du/dt=Au\) and \(u_{k+1}=Au_{k}\), starting from \(u=(9,4)\).
* Find the eigenvalues and eigenvectors of \[A=\begin{bmatrix}0&-i&0\\ i&1&i\\ 0&-i&0\end{bmatrix}.\] What property do you expect for the eigenvectors, and is it true?
* By trying to solve \[\begin{bmatrix}a&b\\ c&d\end{bmatrix}\begin{bmatrix}a&b\\ c&d\end{bmatrix}=\begin{bmatrix}0&1\\ 0&0\end{bmatrix}=A\] show that \(A\) has no square root. Change the diagonal entries of \(A\) to \(4\) and find a square root.
* Find the eigenvalues and eigenvectors of \(A=\begin{bmatrix}0&4\\ \frac{1}{4}&0\end{bmatrix}\).
* Solve \(du/dt=Au\) starting from \(u(0)=(100,100)\).
* If \(v(t)=\) income to stockbrokers and \(w(t)=\) income to client, and they help each other by \(dv/dt=4w\) and \(dw/dt=\frac{1}{4}v\), what does the ratio \(v/w\) approach as \(t\to\infty\)?
* True or false, with reason if true and counterexample if false:
* For every matrix \(A\), there is a solution to \(du/dt=Au\) starting from \(u(0)=(1,\ldots,1)\).
* Every invertible matrix can be diagonalized.
* Every diagonalizable matrix can be inverted.
* Exchanging the rows of a \(2\) by \(2\) matrix reverses the signs of its eigenvalues.
 