12. In three dimensions, \(\lambda_{1}y_{1}^{2}+\lambda_{2}y_{2}^{2}+\lambda_{3}y_{3}^{2}=1\) represents an ellipsoid when all \(\lambda_{i}>0\). Describe all the different kinds of surfaces that appear in the positive semidefinite case when one or more of the eigenvalues is zero.
13. Write down the five conditions for a 3 by 3 matrix to be _negative definite_ (\(-A\) is positive definite) with special attention to condition III: How is \(\det(-A)\) related to \(\det A\)?
14. Decide whether the following matrices are positive definite, negative definite, semidefinite, or indefinite: \[A=\begin{bmatrix}1&2&3\\ 2&5&4\\ 3&4&9\end{bmatrix},\qquad B=\begin{bmatrix}1&2&0&0\\ 2&6&-2&0\\ 0&-2&5&-2\\ 0&0&-2&3\end{bmatrix},\qquad C=-B,\qquad D=A^{-1}.\] Is there a real solution to \(-x^{2}-5y^{2}-9z^{2}-4xy-6xz-8yz=1\)?
15. Suppose \(A\) is symmetric positive definite and \(Q\) is an orthogonal matrix. True or false: 1. \(Q^{\rm T}AQ\) is a diagonal matrix. 2. \(Q^{\rm T}AQ\) is symmetric positive definite. 3. \(Q^{\rm T}AQ\) has the same eigenvalues as \(A\). 4. \(e^{-A}\) is symmetric positive definite.
16. If \(A\) is positive definite and \(a_{11}\) is increased, prove from cofactors that the determinant is increased. Show by example that this can fail if \(A\) is indefinite.
17. From \(A=R^{\rm T}R\). show for positive definite matrices that \(\det A\leq a_{11}a_{22}\cdots a_{mn}\). (The length squared of column \(j\) of \(R\) is \(a_{jj}\). Use determinant \(=\) volume.)
18. (Lyapunov test for stability of \(M\)) Suppose \(AM+M^{\rm H}A=-I\) with positive definite \(A\). If \(Mx=\lambda x\) show that \(\mbox{Re}A<0\). (_Hint_: Multiply the first equation by \(x^{\rm H}\) and \(x\).)
19. Which 3 by 3 symmetric matrices \(A\) produce these functions \(f=x^{\rm T}Ax\)? Why is the first matrix positive definite but not the second one? 1. \(f=2(x_{1}^{2}+x_{2}^{2}+x_{3}^{2}-x_{1}x_{2}-x_{2}x_{3})\). 2. \(f=2(x_{1}^{2}+x_{2}^{2}+x_{3}^{2}-x_{1}x_{2}-x_{1}x_{3}-x_{2}x_{3})\).
20. Compute the three upper left determinants to establish positive definiteness. Verify that their ratios give the second and third pivots.

 