vector \((1,1,-1)\) and automatically contains any multiple \((c,c,-c)\):

\[\textbf{Nullspace is a line}\qquad\begin{bmatrix}1&0&1\\ 5&4&9\\ 2&4&6\end{bmatrix}\begin{bmatrix}c&c&-c\end{bmatrix}=\begin{bmatrix}0\\ 0\\ 0\end{bmatrix}.\]

The nullspace of \(B\) is the line of all points \(x=c\), \(y=c\), \(z=-c\). (The line goes through the origin, as any subspace must.) We want to be able, for any system \(Ax=b\), to find \(C(A)\) and \(N(A)\): all attainable right-hand sides \(b\) and all solutions to \(Ax=0\).

The vectors \(b\) are in the column space and the vectors \(x\) are in the nullspace. We shall compute the dimensions of those subspaces and a convenient set of vectors to generate them. We hope to end up by understanding all _four_ of the subspaces that are intimately related to each other and to \(A\)--the column space of \(A\), the nullspace of \(A\), and their two perpendicular spaces.

Problem Set 2.1
1. Construct a subset of the $x-y$ plane $\mathbf{R}^2$ that is
(a) closed under vector addition and subtraction, but not scalar multiplication.
(b) closed under scalar multiplication but not under vector addition.
Hint: Starting with $u$ and $v$, add and subtract for (a). Try $c u$ and $c v$ for (b).
2. Which of the following subsets of $\mathbf{R}^3$ are actually subspaces?
(a) The plane of vectors $\left(b_1, b_2, b_3\right)$ with first component $b_1=0$.
(b) The plane of vectors $b$ with $b_1=1$.
(c) The vectors $b$ with $b_2 b_3=0$ (this is the union of two subspaces, the plane $b_2=0$ and the plane $b_3=0$ ).
(d) All combinations of two given vectors $(1,1,0)$ and $(2,0,1)$.
(e) The plane of vectors $\left(b_1, b_2, b_3\right)$ that satisfy $b_3-b_2+3 b_1=0$.
3. Describe the column space and the nullspace of the matrices
$$
A=\left[\begin{array}{cc}
1 & -1 \\
0 & 0
\end{array}\right] \quad \text { and } B=\left[\begin{array}{lll}
0 & 0 & 3 \\
1 & 2 & 3
\end{array}\right] \text { and } C=\left[\begin{array}{lll}
0 & 0 & 0 \\
0 & 0 & 0
\end{array}\right] \text {. }
$$
4. What is the smallest subspace of 3 by 3 matrices that contains all symmetric matrices and all lower triangular matrices? What is the largest subspace that is contained in both of those subspaces?
5. Addition and scalar multiplication are required to satisfy these eight rules:
 
