

**25.** Which subspaces are the same for these matrices of different sizes?

\[\text{(a)}\quad\begin{bmatrix}A\end{bmatrix}\quad\text{and}\quad\begin{bmatrix} A\\ A\end{bmatrix}.\qquad\qquad\text{(b)}\quad\begin{bmatrix}A\\ A\end{bmatrix}\quad\text{and}\quad\begin{bmatrix}A&A\\ A&A\end{bmatrix}.\]

Prove that all three matrices have the same rank \(r\).

**26.** If the entries of a 3 by 3 matrix are chosen randomly between 0 and 1, what are the most likely dimensions of the four subspaces? What if the matrix is 3 by 5?

**27.** (Important) \(A\) is an \(m\) by \(n\) matrix of rank \(r\). Suppose there are right-hand sides \(b\) for which \(Ax=b\) has _no solution_.

1. What inequalities (\(<\) or \(\leq\)) must be true between \(m\), \(n\), and \(r\)?
2. How do you know that \(A^{\text{T}}y=0\) has a nonzero solution?

**28.** Construct a matrix with \((1,0,1)\) and \((1,2,0)\) as a basis for its row space and its column space. Why can't this be a basis for the row space and nullspace?

**29.** Without computing \(A\), find bases for the four fundamental subspaces:

\[A=\begin{bmatrix}1&0&0\\ 6&1&0\\ 9&8&1\end{bmatrix}\begin{bmatrix}1&2&3&4\\ 0&1&2&3\\ 0&0&1&2\end{bmatrix}.\]

**30.** If you exchange the first two rows of a matrix \(A\), which of the four subspaces stay the same? If \(y=(1,2,3,4)\) is in the left nullspace of \(A\), write down a vector in the left nullspace of the new matrix.

**31.** Explain why \(v=(1,0,-1)\) cannot be a row of \(A\) and also be in the nullspace.

**32.** Describe the four subspaces of \(\mathbf{R}^{3}\) associated with

\[A=\begin{bmatrix}0&1&0\\ 0&0&1\\ 0&0&0\end{bmatrix}\quad\text{and}\quad I+A=\begin{bmatrix}1&1&0\\ 0&1&1\\ 0&0&1\end{bmatrix}.\]

**33.** (Left nullspace) Add the extra column \(b\) and reduce \(A\) to echelon form:

\[\begin{bmatrix}A&b\end{bmatrix}=\begin{bmatrix}1&2&3&b_{1}\\ 4&5&6&b_{2}\\ 7&8&9&b_{3}\end{bmatrix}\to\begin{bmatrix}1&2&3&b_{1}\\ 0&-3&-6&b_{2}-4b_{1}\\ 0&0&0&b_{3}-2b_{2}+b_{1}\end{bmatrix}.\]

A combination of the rows of \(A\) has produced the zero row. What combination is it? (Look at \(b_{3}-2b_{2}+b_{1}\) on the right-hand side.) Which vectors are in the nullspace of \(A^{\text{T}}\) and which are in the nullspace of \(A\)?