

**1.4**: What is the echelon form \(U\) of \(A\)? \[A=\begin{bmatrix}1&2&0&2&1\\ -1&-2&1&1&0\\ 1&2&-3&-7&-2\end{bmatrix}.\] What are the dimensions of its four fundamental subspaces? **1.5**: Find the rank and the nullspace of \[A=\begin{bmatrix}0&0&1\\ 0&0&1\\ 1&1&1\end{bmatrix}\qquad\text{and}\qquad B=\begin{bmatrix}0&0&1&2\\ 0&0&1&2\\ 1&1&1&0\end{bmatrix}.\]
**1.6**: Find bases for the four fundamental subspaces associated with \[A=\begin{bmatrix}1&2\\ 3&6\end{bmatrix},\qquad B=\begin{bmatrix}0&0\\ 1&2\end{bmatrix},\qquad C=\begin{bmatrix}1&1&0&0\\ 0&1&0&1\end{bmatrix}.\]
**1.7**: What is the most general solution to \(u+v+w=1\), \(u-w=2\)? **1.8**: 1. Construct a matrix whose nullspace contains the vector \(x=(1,1,2)\). 2. Construct a matrix whose left nullspace contains \(y=(1,5)\). 3. Construct a matrix whose column space is spanned by \((1,1,2)\) and whose row space is spanned by \((1,5)\). 4. If you are given any three vectors in \(\mathbf{R}^{6}\) and any three vectors in \(\mathbf{R}^{5}\), is there a 6 by 5 matrix whose column space is spanned by the first three and whose row space is spanned by the second three? **1.9**: In the vector space of 2 by 2 matrices, 1. is the set of rank 1 matrices a subspace? 2. what subspace is spanned by the permutation matrices? 3. what subspace is spanned by the positive matrices (all \(a_{ij}>0\))? 4. what subspace is spanned by the invertible matrices?
**1.10**: Invent a vector space that contains all linear transformations from \(\mathbf{R}^{n}\) to \(\mathbf{R}^{n}\). You have to decide on a rule for addition. What is its dimension?
**1.11**: 1. Find the rank of \(A\), and give a basis for its nullspace. \[A=LU=\begin{bmatrix}1&&&&\\ 2&1&&\\ 2&1&2&\\ 3&2&4&1\end{bmatrix}\begin{bmatrix}1&2&0&1&2&1\\ 0&0&2&2&0&0\\ 0&0&0&0&0&1\\ 0&0&0&0&0&0\end{bmatrix}.\]