* If eigenvectors \(x\) and \(y\) correspond to distinct eigenvalues, then \(x^{\rm H}y=0\).
* If \(K\) is a skew-symmetric matrix, show that \(Q=(I-K)(I+K)^{-1}\) is an orthogonal matrix. Find \(Q\) if \(K=\left[\begin{smallmatrix}0&2\\ -2&0\end{smallmatrix}\right]\).
* If \(K^{\rm H}=-K\) (skew-Hermitian), the eigenvalues are imaginary and the eigenvectors are orthogonal.
* How do you know that \(K-I\) is invertible?
* How do you know that \(K=U\Lambda U^{\rm H}\) for a unitary \(U\)?
* Why is \(e^{\Lambda t}\) unitary?
* Why is \(e^{Kt}\) unitary?
* If \(M\) is the diagonal matrix with entries \(d\), \(d^{2}\), \(d^{3}\), what is \(M^{-1}AM\)? What are its eigenvalues in the following case? \[A=\left[\begin{matrix}1&1&1\\ 1&1&1\\ 1&1&1\end{matrix}\right].\]
* If \(A^{2}=-I\), what are the eigenvalues of \(A\)? If \(A\) is a real \(n\) by \(n\) matrix show that \(n\) must be even, and give an example.
* If \(Ax=\lambda_{1}x\) and \(A^{\rm T}y=\lambda_{2}y\) (all real), show that \(x^{\rm T}y=0\).
* A variation on the Fourier matrix is the "sine matrix": \[S=\frac{1}{\sqrt{2}}\left[\begin{matrix}\sin\theta&\sin 2\theta&\sin 3 \theta\\ \sin 2\theta&\sin 4\theta&\sin 6\theta\\ \sin 3\theta&\sin 6\theta&\sin 9\theta\end{matrix}\right]\quad\mbox{with} \quad\theta=\frac{\pi}{4}.\] Verify that \(S^{\rm T}=S^{-1}\). (The columns are the eigenvectors of the tridiagonal \(-1\), \(2\), \(-1\) matrix.)
* Find a nonzero matrix \(N\) such that \(N^{3}=0\).
* If \(Nx=\lambda x\), show that \(\lambda\) must be zero.
* Prove that \(N\) (called a "nilpotent" matrix) cannot be symmetric.
* Find the matrix \(P=aa^{\rm T}/a^{\rm T}a\) that projects any vector onto the line through \(a=(2,1,2)\).
* What is the only nonzero eigenvalue of \(P\), and what is the corresponding eigenvector?
* Solve \(u_{k+1}=Pu_{k}\), starting from \(u_{0}=(9,9,0)\).
* Suppose the first row of \(A\) is \(7\), \(6\) and its eigenvalues are \(i\), \(-i\). Find \(A\).

 