15. If the columns of \(A\) are linearly independent (\(A\) is \(m\) by \(n\)), then the rank is , the nullspace is , the row space is , and there exists a -inverse.
16. (\(A\)_paradox_) Suppose \(A\) has a right-inverse \(B\). Then \(AB=I\) leads to \(A^{\mathrm{T}}AB=A^{\mathrm{T}}\) or \(B(A^{\mathrm{T}}A)^{-1}A^{\mathrm{T}}\). But that satisfies \(BA=I\); it is a _left_-inverse. Which step is not justified?
17. Find a matrix \(A\) that has \(\mathbf{V}\) as its row space, and a matrix \(B\) that has \(\mathbf{V}\) as its nullspace, if \(\mathbf{V}\) is the subspace spanned by \[\begin{bmatrix}1\\ 1\\ 0\end{bmatrix},\qquad\begin{bmatrix}1\\ 2\\ 0\end{bmatrix},\qquad\begin{bmatrix}1\\ 5\\ 0\end{bmatrix}.\]
18. Find a basis for each of the four subspaces of \[A=\begin{bmatrix}0&1&2&3&4\\ 0&1&2&4&6\\ 0&0&0&1&2\end{bmatrix}=\begin{bmatrix}1&0&0\\ 1&1&0\\ 0&1&1\end{bmatrix}\begin{bmatrix}0&1&2&3&4\\ 0&0&0&1&2\\ 0&0&0&0&0\end{bmatrix}.\]
19. If \(A\) has the same four fundamental subspaces as \(B\), does \(A=cB\)?
20. 1. If a 7 by 9 matrix has rank 5, what are the dimensions of the four subspaces? What is the sum of all four dimensions? 2. If a 3 by 4 matrix has rank 3, what are its column space and left nullspace?
21. Construct a matrix with the required property, or explain why you can't. 1. Column space contains \(\begin{bmatrix}1\\ 0\end{bmatrix}\), \(\begin{bmatrix}0\\ 0\end{bmatrix}\), row space contains \(\begin{bmatrix}1\\ 2\end{bmatrix}\), \(\begin{bmatrix}2\\ 5\end{bmatrix}\). 2. Column space has basis \(\begin{bmatrix}1\\ 2\end{bmatrix}\), nullspace has basis \(\begin{bmatrix}3\\ 2\end{bmatrix}\). 3. Dimension of nullspace \(=1+\) dimension of left nullspace. 4. Left nullspace contains \(\begin{bmatrix}1\\ 3\end{bmatrix}\), row space contains \(\begin{bmatrix}3\\ 1\end{bmatrix}\). 5. Row space \(=\) column space, nullspace \(\neq\) left nullspace.
22. Without elimination, find dimensions and bases for the four subspaces for \[A=\begin{bmatrix}0&3&3&3\\ 0&0&0&0\\ 0&1&0&1\end{bmatrix}\quad\text{and}\quad B=\begin{bmatrix}1&1\\ 4&4\\ 5&5\end{bmatrix}.\]
23. Suppose the 3 by 3 matrix \(A\) is invertible. Write bases for the four subspaces for \(A\), and also for the 3 by 6 matrix \(B=[A\;\;A]\).
24. What are the dimensions of the four subspaces for \(A\), \(B\), and \(C\), if \(I\) is the 3 by 3 identity matrix and 0 is the 3 by 2 zero matrix? 