

**9.**: Show that the trace equals the sum of the eigenvalues, in two steps. First, find the coefficient of \((-\lambda)^{n-1}\) on the right side of equation (16). Next, find all the terms in

\[\det(A-\lambda I)=\det\begin{bmatrix}a_{11}-\lambda&a_{12}&\cdots&a_{1n}\\ a_{21}&a_{22}-\lambda&\cdots&a_{2n}\\ \vdots&\vdots&&\vdots\\ a_{n1}&a_{n2}&\cdots&a_{nn}-\lambda\end{bmatrix}\]

that involve \((-\lambda)^{n-1}\). They all come from the main diagonal! Find that coefficient of \((-\lambda)^{n-1}\) and compare.
**10.**:
1. Construct 2 by 2 matrices such that the eigenvalues of \(AB\) are not the products of the eigenvalues of \(A\) and \(B\), and the eigenvalues of \(A+B\) are not the sums of the individual eigenvalues.
2. Verify, however, that the sum of the eigenvalues of \(A+B\) equals the sum of all the individual eigenvalues of \(A\) and \(B\), and similarly for products. Why is this true?
**11.**: _The eigenvalues of \(A\) equal the eigenvalues of \(A^{\rm T}\)._ This is because \(\det(A-\lambda I)\) equals \(\det(A^{\rm T}-\lambda I)\). That is true because . Show by an example that the eigenvectors of \(A\) and \(A^{\rm T}\) are _not_ the same.
**12.**: Find the eigenvalues and eigenvectors of

\[A=\begin{bmatrix}3&4\\ 4&-3\end{bmatrix}\qquad\text{and}\qquad A=\begin{bmatrix}a&b\\ b&a\end{bmatrix}.\]
**13.**: If \(B\) has eigenvalues 1, 2, 3, \(C\) has eigenvalues 4, 5, 6, and \(D\) has eigenvalues 7, 8, 9, what are the eigenvalues of the 6 by 6 matrix \(A=\begin{bmatrix}B&C\\ 0&D\end{bmatrix}\)?
**14.**: Find the rank and all four eigenvalues for both the matrix of ones and the checker board matrix:

\[A=\begin{bmatrix}1&1&1&1\\ 1&1&1&1\\ 1&1&1&1\\ 1&1&1&1\end{bmatrix}\qquad\text{and}\qquad C=\begin{bmatrix}0&1&0&1\\ 1&0&1&0\\ 0&1&0&1\\ 1&0&1&0\end{bmatrix}.\]

Which eigenvectors correspond to nonzero eigenvalues?
**15.**: What are the rank and eigenvalues when \(A\) and \(C\) in the previous exercise are \(n\) by \(n\)? Remember that the eigenvalue \(\lambda=0\) is repeated \(n-r\) times.
**16.**: If \(A\) is the 4 by 4 matrix of ones, find the eigenvalues and the determinant of \(A-I\).

