2. The first 3 rows of \(U\) are a basis for the row space of \(A\)--true or false? Columns 1, 3, 6 of \(U\) are a basis for the column space of \(A\)--true or false? The four rows of \(A\) are a basis for the row space of \(A\)--true or false? 3. Find as many linearly independent vectors \(b\) as possible for which \(Ax=b\) has a solution. 4. In elimination on \(A\), what multiple of the third row is subtracted to knock out the fourth row?

**1.12**: If \(A\) is an \(n\) by \(n-1\) matrix, and its rank is \(n-2\), what is the dimension of its nullspace?

**1.13**: Use elimination to find the triangular factors in \(A=LU\), if

\[A=\begin{bmatrix}a&a&a&a\\ a&b&b&b\\ a&b&c&c\\ a&b&c&d\end{bmatrix}.\]

Under what conditions on the numbers \(a\), \(b\), \(c\), \(d\) are the columns linearly independent?

**1.14**: Do the vectors \((1,1,3)\), \((2,3,6)\), and \((1,4,3)\) form a basis for \(\mathbf{R}^{3}\)?

**1.15**: What do you know about \(C(A)\) when the number of solutions to \(Ax=b\) is

1. 0 or 1, depending on \(b\).
2. \(\infty\), independent of \(b\).
3. 0 or \(\infty\), depending on \(b\).
4. 1, regardless of \(b\).

**1.16**: In the previous exercise, how is \(r\) related to \(m\) and \(n\) in each example?

**1.17**: If \(x\) is a vector in \(\mathbf{R}^{n}\), and \(x^{\mathrm{T}}y=0\) for every \(y\), prove that \(x=0\).

**1.18**: If \(A\) is an \(n\) by \(n\) matrix such that \(A^{2}=A\) and \(\mathrm{rank}A=n\), prove that \(A=I\).

**1.19**: What subspace of 3 by 3 matrices is spanned by the elementary matrices \(E_{ij}\), with 1s on the diagonal and at most one nonzero entry below?

**1.20**: How many 5 by 5 permutation matrices are there? Are they linearly independent? Do they span the space of all 5 by 5 matrices? No need to write them all down.

**1.21**: What is the rank of the \(n\) by \(n\) matrix with every entry equal to 1? How about the "checkerboard matrix," with \(a_{ij}=0\) when \(i+j\) is even, \(a_{ij}=1\) when \(i+j\) is odd? 