

**31.** Compute the determinants \(S_{1},S_{2},S_{3}\) of these 1, 3, 1 tridiagonal matrices:

\[S_{1}\left|3\right|\qquad S_{2}=\left|\begin{matrix}3&1\\ 1&3\end{matrix}\right|\qquad S_{3}=\left|\begin{matrix}3&1&0\\ 1&3&1\\ 0&1&3\end{matrix}\right|.\]

Make a Fibonacci guess for \(S_{4}\) and verify that you are right.
**32.** Cofactors of those 1, 3, 1 matrices give \(S_{n}=3S_{n-1}-S_{n-2}\). Challenge: _Show that \(S_{n}\) is the Fibonacci number \(F_{2n+2}\) by proving \(F_{2n+2}=3F_{2n}-F_{2n-2}\)._ Keep using Fibonacci's rule \(F_{k}=F_{k-1}+F_{k-2}\).
**33.** Change 3 to 2 in the upper left corner of the matrices in Problem 32. Why does that subtract \(S_{n-1}\) from the determinant \(S_{n}\)? Show that the determinants become the Fibonacci numbers 2, 5, 13 (always \(F_{2n+1}\)).

**Problems 34-36 are about block matrices and block determinants.**

**34.** With 2 by 2 blocks, you cannot always use block determinants!

\[\left|\begin{matrix}A&B\\ 0&D\end{matrix}\right|=|A||D|\qquad\text{but}\qquad\left|\begin{matrix}A&B\\ C&D\end{matrix}\right|\neq|A||D|-|C||B|.\]

1. Why is the first statement true? Somehow \(B\) doesn't enter.
2. Show by example that equality fails (as shown) when \(C\) enters.
3. Show by example that the answer \(\det(AD-CB)\) is also wrong.
**35.** With block multiplication, \(A=LU\) has \(A_{k}=L_{k}U_{k}\) in the upper left corner:

\[A=\left[\begin{matrix}A_{k}&*\\ *&*\end{matrix}\right]=\left[\begin{matrix}L_{k}&0\\ *&*\end{matrix}\right]\left[\begin{matrix}U_{k}&*\\ 0&*\end{matrix}\right].\]

1. Suppose the first three pivots of \(A\) are 2, 3, \(-1\). What are the determinants of \(L_{1}\), \(L_{2}\), \(L_{3}\) (with diagonal 1s), \(U_{1}\), \(U_{2}\), \(U_{3}\), and \(A_{1}\), \(A_{2}\), \(A_{3}\)?
2. If \(A_{1}\), \(A_{2}\), \(A_{3}\) have determinants 5, 6, 7, find the three pivots.
**36.** Block elimination subtracts \(CA^{-1}\) times the first row \([A\;\;B]\) from the second row \([C\;\;D]\). This leaves the _Schur complement_\(D-CA^{-1}B\) in the corner:

\[\left[\begin{matrix}I&0\\ -CA^{-1}&I\end{matrix}\right]\left[\begin{matrix}A&B\\ C&D\end{matrix}\right]=\left[\begin{matrix}A&B\\ 0&D-CA^{-1}B\end{matrix}\right].\]

Take determinants of these matrices to prove correct rules for square blocks:

\[\left|\begin{matrix}A&B\\ C&D\end{matrix}\right|=|A|\left|\begin{matrix}D-CA^{-1}B\\ \text{if }A^{-1}\text{ exists}\end{matrix}\right|=|AD-CB|\,.\]