3. \(P\) exchanges rows 1 and 2, then rows 2 and 3.
23. In Problem 22, applying \(E_{21}\) and then \(E_{32}\) to the column \(b=(1,0,0)\) gives \(E_{32}E_{21}b=\). Applying \(E_{32}\) before \(E_{21}\) gives \(E_{21}E_{32}b=\). When \(E_{32}\) comes first, row feels no effect from row.
24. Which three matrices \(E_{21}\), \(E_{31}\), \(E_{32}\) put \(A\) into triangular form \(U\)? \[A=\begin{bmatrix}1&1&0\\ 4&6&1\\ -2&2&0\end{bmatrix}\quad\text{and}\quad E_{32}E_{31}E_{21}A=U.\] Multiply those \(E\)'s to get one matrix \(M\) that does elimination: \(MA=U\).
25. Suppose \(a_{33}=7\) and the third pivot is 5. If you change \(a_{33}\) to 11, the third pivot is . If you change \(a_{33}\) to , there is zero in the pivot position.
26. If every column of \(A\) is a multiple of \((1,1,1)\), then \(Ax\) is always a multiple of \((1,1,1)\). Do a 3 by 3 example. How many pivots are produced by elimination?
27. What matrix \(E_{31}\) subtracts 7 times row 1 from row 3? To reverse that step, \(R_{31}\) should . If you change \(a_{33}\) to , there is zero in the pivot position.
28. If every column of \(A\) is a multiple of \((1,1,1)\), then \(Ax\) is always a multiple of \((1,1,1)\). Do a 3 by 3 example. How many pivots are produced by elimination?
27. What matrix \(E_{31}\) subtracts 7 times row 1 from row 3? To reverse that step, \(R_{31}\) should . If you change \(a_{33}\) to , there is zero in the pivot position.
28. If every column of \(A\) is a multiple of \((1,1,1)\), then \(Ax\) is always a multiple of \((1,1,1)\). Do a 3 by 3 example. How many pivots are produced by elimination?
29. If every column of \(A\) is a multiple of \((1,1,1)\), then \(Ax\) is always a multiple of \((1,1,1)\). Do a 3 by 3 example. How many pivots are produced by elimination?
30. If every column of \(A\) is a multiple of \ 