timated the final roundoff error. It was Wilkinson who found the right way to answer the question, and his books are now classics.

Two simple examples will illustrate three important points about roundoff error. The examples are

\[\mbox{\bf Ill-conditioned}\quad A=\begin{bmatrix}1.&1.\\ 1.&1.0001\end{bmatrix}\qquad\mbox{\bf Well-conditioned}\quad B=\begin{bmatrix}.0001&1. \\ 1.&1.\end{bmatrix}.\]

\(A\) is nearly singular whereas \(B\) is far from singular. If we slightly change the last entry of \(A\) to \(a_{22}=1\), it _is_ singular. Consider two very close right-hand sides \(b\):

\[\begin{array}{ccccccc}u&+&v&=&2\\ u&+&1.0001v&=&2\end{array}\qquad\mbox{and}\qquad\begin{array}{ccccccc}u&+&v&=& 2\\ u&+&1.0001v&=&2.0001\end{array}\]

The solution to the first is \(u=2\), \(v=0\). The solution to the second is \(u=v=1\). _A change in the fifth digit of \(b\) was amplified to a change in the first digit of the solution. No numerical method can avoid this sensitivity to small perturbations_. The ill-conditioning can be shifted from one place to another, but it cannot be removed. The true solution is very sensitive, and the computed solution cannot be less so.

The second point is as follows.

## 10 Even a well-conditioned matrix like \(B\) can be ruined by a poor algorithm.

We regret to say that for the matrix \(B\), direct Gaussian elimination is a poor algorithm. Suppose \(.0001\) is accepted as the first pivot. Then 10,000 times the first row is subtracted from the second. The lower right entry becomes \(-9999\), but roundoff to three places would give \(-10\),000. Every trace of the entry 1 would disappear:

\[\begin{array}{ccccccc}\mbox{\bf Elimination on }B&.0001u+v&=&1\\ \mbox{\bf with small pivot}&u+v&=&2\end{array}\longrightarrow\begin{array}{ ccc}.0001u+v&=&1\\ -9999v&=&-9998.\end{array}\]

Roundoff will produce \(-10,000v=-10,000\), or \(v=1\). This is correct to three decimal places. Back-substitution with the right \(v=.9999\) would leave \(u=1\):

\[\mbox{\bf Correct result}\qquad.0001u+.9999=1,\quad\mbox{or}\quad u=1.\]

Instead, accepting \(v=1\), which is wrong only in the fourth place, we obtain \(u=0\):

\[\mbox{\bf Wrong result}\qquad.0001u+1=1,\quad\mbox{or}\quad u=0.\]

The computed \(u\) is completely mistaken. \(B\) is well-conditioned but elimination is violently unstable. \(L\), \(D\), and \(U\) are completely out of scale with \(B\):

\[B=\begin{bmatrix}1&0\\ 10,000&1\end{bmatrix}\begin{bmatrix}.0001&0\\ 0&-9999\end{bmatrix}\begin{bmatrix}1&10,000\\ 0&1\end{bmatrix}.\]

The small pivot \(.0001\) brought instability, and the remedy is clear--_exchange rows_.

 