1. \(x+y=y+x\).
2. \(x+(y+z)=(x+y)+z\).
3. There is a unique "zero vector" such that \(x+0=x\) for all \(x\).
4. For each \(x\) there is a unique vector \(-x\) such that \(x+(-x)=0\).
5. \(1x=x\).
6. \((c_{1}c_{2})x=c_{1}(c_{2}x)\).
7. \(c(x+y)=cx+cy\).
8. \((c_{1}+c_{2})x=c_{1}x+c_{2}x\).

1. Suppose addition in \(\mathbf{R}^{2}\) adds an extra 1 to each component, so that \((3,1)+(5,0)\) equals \((9,2)\) instead of \((8,1)\). With scalar multiplication unchanged, which rules are broken?
2. Show that the set of all positive real numbers, with \(x+y\) and \(cx\) redefined to equal the usual \(xy\) and \(x^{c}\), is a vector space. What is the "zero vector"?
3. Suppose \((x_{1},x_{2})+(y_{1},y_{2})\) is defined to be \((x_{1}+y_{2},x_{2}+y_{1})\). With the usual \(cx=(cx_{1},cx_{2})\), which of the eight conditions are not satisfied?

**6.**: Let \(\mathbf{P}\) be the plane in 3-space with equation \(x+2y+z=6\). What is the equation of the plane \(\mathbf{P}_{0}\) through the origin parallel to \(\mathbf{P}\)? Are \(\mathbf{P}\) and \(\mathbf{P}_{0}\) subspaces of \(\mathbf{R}^{3}\)?

**7.**: Which of the following are subspaces of \(\mathbf{R}^{\infty}\)?

1. All sequences like \((1,0,1,0,\ldots)\) that include infinitely many zeros.
2. All sequences \((x1,x2,\ldots)\) with \(x_{j}=0\) from some point onward.
3. All decreasing sequences: \(x_{j+1}\leq x_{j}\) for each \(j\).
4. All convergent sequences: the \(x_{j}\) have a limit as \(j\to\infty\).
5. All arithmetic progressions: \(x_{j+1}-x_{j}\) is the same for all \(j\).
6. All geometric progressions \((x_{1},kx_{1},k^{2}x_{1},\ldots)\) allowing all \(k\) and \(x_{1}\).

**8.**: Which of the following descriptions are correct? The solutions \(x\) of

\[Ax=\begin{bmatrix}1&1&1\\ 1&0&2\end{bmatrix}\begin{bmatrix}x_{1}\\ x_{2}\\ x_{3}\end{bmatrix}=\begin{bmatrix}0\\ 0\end{bmatrix}\]

form

1. a plane.
2. a line.
3. a point.
4. a subspace.

 