

**6.**: When \(b=(2,5,7)\), find a solution \((u,v,w)\) to equation (4) different from the solution \((1,0,1)\) mentioned in the text.
**7.**: Give two more right-hand sides in addition to \(b=(2,5,7)\) for which equation (4) can be solved. Give two more right-hand sides in addition to \(b=(2,5,6)\) for which it cannot be solved.
**8.**: Explain why the system

\[\begin{array}{ccccccc}u&+&v&+&w&=&2\\ u&+&2v&+&3w&=&1\\ &&v&+&2w&=&0\end{array}\]

is singular by finding a combination of the three equations that adds up to \(0=1\). What value should replace the last zero on the right side to allow the equations to have solutions--and what is one of the solutions?
**9.**: The column picture for the previous exercise (singular system) is

\[u\begin{bmatrix}1\\ 1\\ 0\end{bmatrix}+v\begin{bmatrix}1\\ 2\\ 1\end{bmatrix}+w\begin{bmatrix}1\\ 3\\ 2\end{bmatrix}=b.\]

Show that the three columns on the left lie in the same plane by expressing the third column as a combination of the first two. What are all the solutions \((u,v,w)\) if \(b\) is the zero vector \((0,0,0)\)?
**10.**: (Recommended) Under what condition on \(y_{1}\), \(y_{2}\), \(y_{3}\) do the points \((0,y_{1})\), \((1,y_{2})\), \((2,y_{3})\) lie on a straight line?
**11.**: These equations are certain to have the solution \(x=y=0\). For which values of \(a\) is there a whole line of solutions?

\[\begin{array}{ccccccc}ax&+&2y&=&0\\ 2x&+&ay&=&0\end{array}\]
**12.**: Starting with \(x+4y=7\), find the equation for the parallel line through \(x=0\), \(y=0\). Find the equation of another line that meets the first at \(x=3\), \(y=1\).

**Problems 13-15 are a review of the row and column pictures.**
**13.**: Draw the two pictures in two planes for the equations \(x-2y=0\), \(x+y=6\).
**14.**: For two linear equations in three unknowns \(x\), \(y\), \(z\), the row picture will show (2 or 3) (lines or planes) in (two or three)-dimensional space. The column picture is in (two or three)-dimensional space. The solutions normally lie on a