15. Cramer's Rule breaks down when \(\det A=0\). Example (a) has no solution, whereas (b) has infinitely many. What are the ratios \(x_{j}=\det B_{j}/\det A\)? (a) \(\begin{array}{ll}2x_{1}+3x_{2}=1\\ 4x_{1}+6x_{2}=1.\end{array}\) (parallel lines) (b) \(\begin{array}{ll}2x_{1}+3x_{2}=1\\ 4x_{1}+6x_{2}=2.\end{array}\) (same line)
16. _Quick proof of Cramer's rule_. The determinant is a linear function of column 1. It is zero if two columns are equal. When \(\boldsymbol{b}=A\boldsymbol{x}=x_{1}\boldsymbol{a_{1}}+x_{2}\boldsymbol{a_{2 }}+x_{3}\boldsymbol{a_{3}}\) goes into column 1 to produce \(B_{1}\), the determinant is (a) What formula for \(x_{1}\) comes from left side \(=\) right side? (b) What steps lead to the middle equation?
17. If the right side \(b\) is the _last column_ of \(A\), solve the 3 by 3 system \(Ax=b\). Explain how each determinant in Cramer's Rule leads to your solution \(x\). Problems 18-26 are about \(A^{-1}=C^{\mathrm{T}}/\det A\). Remember to transpose \(C\).
18. Find \(A^{-1}\) from the cofactor formula \(C^{\mathrm{T}}/\det A\). Use symmetry in part (b): (a) \(\begin{array}{ll}A=\begin{bmatrix}1&2&0\\ 0&3&0\\ 0&4&1\end{bmatrix}.\end{array}\) (b) \(\begin{array}{ll}A=\begin{bmatrix}2&-1&0\\ -1&2&-1\\ 0&-1&2\end{bmatrix}.\end{array}\)
19. If all the cofactors are zero, how do you know that \(A\) has no inverse? If none of the cofactors are zero, is \(A\) sure to be invertible?
20. Find the cofactors of \(A\) and multiply \(AC^{\mathrm{T}}\) to find \(\det A\): \[A=\begin{bmatrix}1&1&4\\ 1&2&2\\ 1&2&5\end{bmatrix},\qquad C=\begin{bmatrix}6&-3&0\\ .&.&.\\ .&.&.\end{bmatrix},\quad\text{and}\quad AC^{\mathrm{T}}=\underline{\phantom{ \underline{\underline{\underline{\underline{\underline{\underline{\underline{ \underline{\underline{\underline{\underline{\underline{\underline{\underline{ \underline{\underline{ \underline{         }}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}}........\.\...\..................................

If you change that you change that corner entry from \ \ 1000000000}\}\.\.\.\.\.\.\.\.\.\.\.\.\.\.\.\.\.\.\.\.\ 