

**9.**: The quadratic \(f(x_{1},x_{2})=3(x_{1}+2x_{2})^{2}+4x_{2}^{2}\) is positive. Find its matrix \(A\), factor it into \(LDL^{\rm T}\), and connect the entries in \(D\) and \(L\) to 3, 2, 4 in \(f\).
**10.**: If \(R=[\begin{smallmatrix}p&q\\ q&r\end{smallmatrix}]\), write out \(R^{2}\) and check that it is positive definite unless \(R\) is singular.
**11.**:
1. If \(A=\left[\begin{smallmatrix}a&b\\ b&c\end{smallmatrix}\right]\) is Hermitian (_complex_\(b\)), find its pivots and determinant. 2. Complete the square for \(x^{\rm H}Ax\). Now \(x^{\rm H}=[\overline{x}_{1}\ \ \overline{x}_{2}]\) can be complex \[a|x_{1}|^{2}+2{\rm Re}b\overline{x}_{1}x_{2}+c|x_{2}|^{2}=a|x_{1}+(b/a)x_{2}|^ {2}+\underline{\phantom{\phantom{\{\{\{\{\{\{\{\{\{\{\{\{\{\{\{\{\{\{\{ }}}}}}}}}}}}}}}}|x_{2}|^{2}.\] 3. Show that \(a>0\) and \(ac>|b|^{2}\) ensure that \(A\) is positive definite. 4. Are the matrices \(\left[\begin{smallmatrix}1&1\\ 1-i&2\end{smallmatrix}+i\right]\) and \(\left[\begin{smallmatrix}3&4+i\\ 4-i&6\end{smallmatrix}\right]\) positive definite?
**12.**: Decide whether \(F=x^{2}y^{2}-2x-2y\) has a minimum at the point \(x=y=1\) (after showing that the first derivatives are zero at that point).
**13.**: Under what conditions on \(a\), \(b\), \(c\) is \(ax^{2}+2bxy+cy^{2}>x^{2}+y^{2}\) for all \(x\), \(y\)?

**Problems 14-18 are about tests for positive definiteness.**

**14.**: Which of \(A_{1},A_{2},A_{3},A_{4}\) has two positive eigenvalues? Test \(a>0\) and \(ac>b^{2}\), don't compute the eigenvalues. Find an \(x\) so that \(x^{\rm T}A_{1}x<0\).

\[A_{1}=\begin{bmatrix}5&6\\ 6&7\end{bmatrix}\qquad A_{2}=\begin{bmatrix}-1&-2\\ -2&-5\end{bmatrix}\qquad A_{3}=\begin{bmatrix}1&10\\ 10&100\end{bmatrix}\qquad A_{4}=\begin{bmatrix}1&10\\ 10&101\end{bmatrix}.\]
**15.**: What is the quadratic \(f=ax^{2}+2bxy+cy^{2}\) for each of these matrices? Complete the square to write \(f\) as a sum of one or two squares \(d_{1}(\qquad)^{2}+d_{2}(\qquad)^{2}\).

\[A=\begin{bmatrix}1&2\\ 2&9\end{bmatrix}\qquad\text{ and }\qquad A=\begin{bmatrix}1&3\\ 3&9\end{bmatrix}.\]
**16.**: Show that \(f(x,y)=x^{2}+4xy+3y^{2}\) does not have a minimum at \((0,0)\) even though it has positive coefficients. Write \(f\) as a _difference_ of squares and find a point \((x,y)\) where \(f\) is negative.
**17.**: (_Important_) If \(A\) has _independent columns_, then \(A^{\rm T}A\) is square and symmetric and invertible (Section 4.2). _Rewrite \(x^{\rm T}A^{\rm T}Ax\) to show why it is positive except when \(x=0\)_. Then \(A^{\rm T}A\) is positive definite.
**18.**: Test to see if \(A^{\rm T}A\) is positive definite in each case:

\[A=\begin{bmatrix}1&2\\ 0&3\end{bmatrix},\qquad A=\begin{bmatrix}1&1\\ 1&2\\ 2&1\end{bmatrix},\quad\text{and}\quad A=\begin{bmatrix}1&1&2\\ 1&2&1\end{bmatrix}.\]