

**6.**: Choose three independent columns of \(U\). Then make two other choices. Do the same for \(A\). You have found bases for which spaces?

\[U=\begin{bmatrix}2&3&4&1\\ 0&6&7&0\\ 0&0&0&9\\ 0&0&0&0\end{bmatrix}\quad\text{and}\quad A=\begin{bmatrix}2&3&4&1\\ 0&6&7&0\\ 0&0&0&9\\ 4&6&8&2\end{bmatrix}.\]
**7.**: If \(w_{1}\), \(w_{2}\), \(w_{3}\) are independent vectors, show that the differences \(v_{1}=w_{2}-w_{3}\), \(v_{2}=w_{1}-w_{3}\), and \(v_{3}=w_{1}-w_{2}\) are _dependent_. Find a combination of the \(v\)'s that gives zero.
**8.**: If \(w_{1}\), \(w_{2}\), \(w_{3}\) are independent vectors, show that the sums \(v_{1}=w_{2}+w_{3}\), \(v_{2}=w_{1}+w_{3}\), and \(v_{3}=w_{1}+w_{2}\) are _independent_. (Write \(c_{1}v_{1}+c_{2}v_{2}+c_{3}v_{3}=0\) in terms of the \(w\)'s. Find and solve equations for the \(c\)'s.)
**9.**: Suppose \(v_{1}\), \(v_{2}\), \(v_{3}\), \(v_{4}\) are vectors in \(\mathbf{R}^{3}\).

1. These four vectors are dependent because .
2. The two vectors \(v_{1}\) and \(v_{2}\) will be dependent if .
3. The vectors \(v_{1}\) and \((0,0,0)\) are dependent because .
**10.**: Find two independent vectors on the plane \(x+2y-3z-t=0\) in \(\mathbf{R}^{4}\). Then find three independent vectors. Why not four? This plane is the nullspace of what matrix?
**Problems 11-18 are about the space _spanned_ by a set of vectors. Take all linear combinations of the vectors
**11.**: Describe the subspace of \(\mathbf{R}^{3}\) (is it a line or a plane or \(\mathbf{R}^{3}\)?) spanned by

1. the two vectors \((1,1,-1)\) and \((-1,-1,1)\).
2. the three vectors \((0,1,1)\) and \((1,1,0)\) and \((0,0,0)\).
3. the columns of a 3 by 5 echelon matrix with 2 pivots.
4. all vectors with positive components.
**12.**: The vector \(b\) is in the subspace spanned by the columns of \(A\) when there is a solution to . The vector \(c\) is in the row space of \(A\) when there is a solution to . _True or false_: If the zero vector is in the row space, the rows are dependent.
**13.**: Find the dimensions of (a) the column space of \(A\), (b) the column space of \(U\), (c) the row space of \(A\), (d) the row space of \(U\). Which two of the spaces are the same?

\[A=\begin{bmatrix}1&1&0\\ 1&3&1\\ 3&1&-1\end{bmatrix}\quad\text{and}\quad U=\begin{bmatrix}1&1&0\\ 0&2&1\\ 0&0&0\end{bmatrix}.\]