

**28.**: Find the _range_ and _kernel_ (those are new words for the column space and nullspace) of \(T\).

\[\begin{array}{llll}\mbox{(a)}&T(v_{1},v_{2})=(v_{2},v_{1}).&\mbox{(b)}&T(v_{1} ,v_{2},v_{3})=(v_{1},v_{2}).\\ \mbox{(c)}&T(v_{1},v_{2})=(0,0).&\mbox{(d)}&T(v_{1},v_{2})=(v_{1},v_{1}).\end{array}\]
**29.**: A linear transformation from \(\mathbf{V}\) to \(\mathbf{W}\) has an _inverse_ from \(\mathbf{W}\) to \(\mathbf{V}\) when the range is all of \(\mathbf{W}\) and the kernel contains only \(v=0\). Why are these transformations not invertible?

1. \(T(v_{1},v_{2})=(v_{2},v_{2})\qquad\mathbf{W}=\mathbf{R}^{2}\).
2. \(T(v_{1},v_{2})=(v_{1},v_{2},v_{1}+v_{2})\qquad\mathbf{W}=\mathbf{R}^{3}\).
3. \(T(v_{1},v_{2})=v_{1}\qquad\mathbf{W}=\mathbf{R}^{1}\).
**30.**: Suppose a linear \(T\) transforms \((1,1)\) to \((2,2)\) and \((2,0)\) to \((0,0)\). Find \(T(v)\) when (a) \(v=(2,2)\). (b) \(v=(3,1)\). (c) \(v=(-1,1)\). (d) \(v=(a,b)\).
**Problems 31-35 may be harder. The input space \(\mathbf{V}\) contains all 2 by 2 matrices \(M\).**
**31.**: \(M\) is any 2 by 2 matrix and \(A=\left[\begin{smallmatrix}1&2\\ 3&4\end{smallmatrix}\right]\). The linear transformation \(T\) is defined by \(T(M)=AM\). What rules of matrix multiplication show that \(T\) is linear?
**32.**: Suppose \(A=\left[\begin{smallmatrix}1&2\\ 3&6\end{smallmatrix}\right]\). Show that the identity matrix \(I\) is not in the range of \(T\). Find a nonzero matrix \(M\) such that \(T(M)=AM\) is zero.
**33.**: Suppose \(T\) transposes every matrix \(M\). Try to find a matrix \(A\) that gives \(AM=M^{\mathrm{T}}\) for every \(M\). Show that no matrix \(A\) will do it. _To professors:_ Is this a linear transformation that doesn't come from a matrix?
**34.**: The transformation \(T\) that transposes every matrix is definitely linear. Which of these extra properties are true?

1. \(T^{2}=\mbox{identity transformation}\).
2. The kernel of \(T\) is the zero matrix.
3. Every matrix is in the range of \(T\).
4. \(T(M)=-M\) is impossible.
**35.**: Suppose \(T(M)=\left[\begin{smallmatrix}1&0\\ 0&0\end{smallmatrix}\right]\left[\begin{smallmatrix}M\end{smallmatrix}\right] \left[\begin{smallmatrix}0&0\\ 0&1\end{smallmatrix}\right]\). Find a matrix with \(T(M)\neq 0\). Describe all matrices with \(T(M)=0\) (the kernel of \(T\)) and all output matrices \(T(M)\) (the range of \(T\)).

**Problems 36-40 are about changing the basis**

**36.**:
1. What matrix transforms \((1,0)\) into \((2,5)\) and transforms \((0,1)\) to \((1,3)\)?
2. What matrix transforms \((2,5)\) to \((1,0)\) and \((1,3)\) to \((0,1)\)?