

**52.**: Reduce these matrices \(A\) and \(B\) to their ordinary echelon forms \(U\):

\[\text{(a)}\quad A=\begin{bmatrix}1&2&2&4&6\\ 1&2&3&6&9\\ 0&0&1&2&3\end{bmatrix}\qquad\qquad\text{(b)}\quad B=\begin{bmatrix}2&4&2\\ 0&4&4\\ 0&8&8\end{bmatrix}.\]

Find a special solution for each free variable and describe every solution to \(Ax=0\) and \(Bx=0\). Reduce the echelon forms \(U\) to \(R\), and draw a box around the identity matrix in the pivot rows and pivot columns.
**53.**: True or False? (Give reason if true, or counterexample to show it is false.)

1. A square matrix has no free variables.
2. An invertible matrix has no free variables.
3. An \(m\) by \(n\) matrix has no more than \(n\) pivot variables.
4. An \(m\) by \(n\) matrix has no more than \(m\) pivot variables.
**54.**: Is there a 3 by 3 matrix with no zero entries for which \(U=R=I\)?
**55.**: Put as many 1s as possible in a 4 by 7 echelon matrix \(U\) and in a _reduced_ form \(R\) whose pivot columns are 2, 4, 5.
**56.**: Suppose column 4 of a 3 by 5 matrix is all 0s. Then \(x_{4}\) is certainly a variable. The special solution for this variable is the vector \(x=\).
**57.**: Suppose the first and last columns of a 3 by 5 matrix are the same (nonzero). Then is a free variable. Find the special solution for this variable.
**58.**: The equation \(x-3y-z=0\) determines a plane in \(\mathbf{R}^{3}\). What is the matrix \(A\) in this equation? Which are the free variables? The special solutions are \((3,1,0)\) and . The parallel plane \(x-3y-z=12\) contains the particular point \((12,0,0)\). All points on this plane have the following form (fill in the first components):

\[\begin{bmatrix}x\\ y\\ z\end{bmatrix}=\begin{bmatrix}0\\ 0\end{bmatrix}+y\begin{bmatrix}1\\ 0\end{bmatrix}+z\begin{bmatrix}0\\ 1\end{bmatrix}.\]
**59.**: Suppose column 1 \(+\) column 3 \(+\) column 5 \(=0\) in a 4 by 5 matrix with four pivots. Which column is sure to have no pivot (and which variable is free)? What is the special solution? What is the nullspace?
**Problems 60-66 ask for matrices (if possible) with specific properties.**
**60.**: Construct a matrix whose nullspace consists of all combinations of \((2,2,1,0)\) and \((3,1,0,1)\).
**61.**: Construct a matrix whose nullspace consists of all multiples of \((4,3,2,1)\).

