

**Problems 25-28 are about the diagonalizability of \(A\).**

**25.**: True or false: If the eigenvalues of \(A\) are 2, 2, 5, then the matrix is certainly

1. invertible.
2. diagonalizable.
3. not diagonalizable.
**26.**: If the eigenvalues of \(A\) are 1 and 0, write everything you know about the matrices \(A\) and \(A^{2}\).
**27.**: Complete these matrices so that \(\det A=25\). Then \(\text{trace}=10\), and \(\lambda=5\) is repeated! Find an eigcvector with \(Ax=5x\). These matrices will nothe diagonalizabie because there is no second line of eigenvectors.

\[A=\begin{bmatrix}8\\ &2\end{bmatrix},\qquad A=\begin{bmatrix}9&4\\ &1\end{bmatrix},\quad\text{and}\quad A=\begin{bmatrix}10&5\\ -5\end{bmatrix}.\]
**28.**: The matrix \(A=\begin{bmatrix}3&1\\ 0&3\end{bmatrix}\) is not diagonalizable because the rank of \(A-3I\) is . Change one entry to make \(A\) diagonalizable. Which entries could you change?

**Problems 29-33 are about powers of matrices.**

**29.**: \(A^{k}=S\Lambda^{k}S^{-1}\) approaches the zero matrix as \(k\to\infty\) if and only if every \(\lambda\) has absolute value less than . Does \(A^{k}\to 0\) or \(B^{k}\to 0\)?

\[A=\begin{bmatrix}.6&.4\\ .4&.6\end{bmatrix}\qquad\text{and}\qquad B=\begin{bmatrix}.6&.9\\ .1&.6\end{bmatrix}.\]
**30.**: (Recommended) Find \(\Lambda\) and \(S\) to diagonalize \(A\) in Problem 29. What is the limit of \(\Lambda^{k}\) as \(k\to\infty\)? What is the limit of \(S\Lambda^{k}S^{-1}\)? In the columns of this limiting matrix you see the .
**31.**: Find \(\Lambda\) and \(S\) to diagonalize \(B\) in Problem 29. What is \(B^{10}u_{0}\) for these \(u_{0}\)?

\[u_{0}=\begin{bmatrix}3\\ 1\end{bmatrix},\qquad u_{0}=\begin{bmatrix}3\\ -1\end{bmatrix},\quad\text{and}\quad u_{0}=\begin{bmatrix}6\\ 0\end{bmatrix}.\]
**32.**: Diagonalize \(A\) and compute \(S\Lambda^{k}S^{-1}\) to prove this formula for \(A^{k}\):

\[A=\begin{bmatrix}2&1\\ 1&2\end{bmatrix}\qquad\text{has}\qquad A^{k}=\frac{1}{2}\begin{bmatrix}3^{k} +1&3^{k}-1\\ 3^{k}-1&3^{k}+1\end{bmatrix}.\]
**33.**: Diagonalize \(B\) and compute \(S\Lambda^{k}S^{-1}\) to prove this formula for \(B^{k}\):

\[B=\begin{bmatrix}3&1\\ 0&2\end{bmatrix}\qquad\text{has}\qquad B^{k}=\begin{bmatrix}3^{k}&3^{k}-2^{k} \\ 0&2^{k}\end{bmatrix}.\]