\[U_{1}=\frac{1}{2}\,(\,T+\,T^{*})\]

\[U_{2}=\frac{1}{2i}\,(\,T-\,T^{*}).\]

A linear operator \(T\) such that \(T=T^{*}\) is called **self-adjoint** (or **Hermitian**). If \(\mathcal{B}\) is an orthonormal basis for \(V\), then

\[[T^{*}]_{\mathcal{B}}=[T]_{\mathcal{B}}^{*}\]

and so \(T\) is self-adjoint if and only if its matrix in every orthonormal basis is a self-adjoint matrix. Self-adjoint operators are important, not simply because they provide us with some sort of real and imaginary part for the general linear operator, but for the following reasons: (1) Self-adjoint operators have many special properties. For example, for such an operator there is an orthonormal basis of characteristic vectors. (2) Many operators which arise in practice are self-adjoint. We shall consider the special properties of self-adjoint operators later.

### Exercises

**1.** Let \(V\) be the space \(C^{\natural}\), with the standard inner product. Let \(T\) be the linear operator defined by \(T\epsilon_{1}=(1,\,-2)\), \(T\epsilon_{2}=(i,\,-1)\). If \(\alpha=(x_{1},x_{2})\), find \(T^{*}\alpha\).

**2.** Let \(T\) be the linear operator on \(C^{\natural}\) defined by \(T\epsilon_{1}=(1+i,\,2)\), \(T\epsilon_{2}=(i,\,i)\). Using the standard inner product, find the matrix of \(T^{*}\) in the standard ordered basis. Does \(T\) commute with \(T^{*}\)?

**3.** Let \(V\) be \(C^{\natural}\) with the standard inner product. Let \(T\) be the linear operator on \(V\) whose matrix in the standard ordered basis is defined by

\[A_{ik}=i^{\prime+k},\qquad(i^{\natural}=-1).\]

Find a basis for the null space of \(T^{*}\).

**4.** Let \(V\) be a finite-dimensional inner product space and \(T\) a linear operator on \(V\). Show that the range of \(T^{*}\) is the orthogonal complement of the null space of \(T\).

**5.** Let \(V\) be a finite-dimensional inner product space and \(T\) a linear operator on \(V\). If \(T\) is invertible, show that \(T^{*}\) is invertible and \((T^{*})^{-1}=(T^{-1})^{*}\).

**6.** Let \(V\) be an inner product space and \(\beta\), \(\gamma\) fixed vectors in \(V\). Show that \(T\alpha=(\alpha|\beta)\gamma\) defines a linear operator on \(V\). Show that \(T\) has an adjoint, and describe \(T^{*}\) explicitly.

Now suppose \(V\) is \(C^{n}\) with the standard inner product, \(\beta=(y_{1},\,.\,.\,.\,,\,y_{n})\), and \(\gamma=(x_{1},\,.\,.\,.\,,\,x_{n})\). What is the \(j\), \(k\) entry of the matrix of \(T\) in the standard ordered basis? What is the rank of this matrix?

**7.** Show that the product of two self-adjoint operators is self-adjoint if and only if the two operators commute.

 