who feels that this paragraph is much ado about nothing might ask himself whether the vectors \[\alpha_{1} = (e^{\pi/2},1)\] \[\alpha_{2} = (\sqrt[3]{110},1)\] are linearly independent in \(R^{2}\). The elements of a sequence are enumerated in a specific order. A set is a collection of objects, with no specified arrangement or order. Of course, to describe the set we may list its members, and that requires choosing an order. But, the order is not part of the set. The sets \(\{1,\,2,\,3,\,4\}\) and \(\{4,\,3,\,2,\,1\}\) are identical, whereas \(1,\,2,\,3,\,4\) is quite a different sequence from \(4,\,3,\,2,\,1\). The order aspect of sequences has no bearing on questions of independence, dependence, etc., because dependence (as defined) is not affected by the order. The sequence \(\alpha_{n}\), \(\ldots\), \(\alpha_{1}\) is dependent if and only if the sequence \(\alpha_{1}\), \(\ldots\), \(\alpha_{n}\) is dependent. In the next section, order will be important.

### Exercises

**1.** Prove that if two vectors are linearly dependent, one of them is a scalar multiple of the other. **2.** Are the vectors \[\begin{array}{l}\alpha_{1}=(1,1,2,4),\quad\ \alpha_{2}=(2,-1,-5,2)\\ \alpha_{3}=(1,-1,-4,0),\quad\ \alpha_{4}=(2,1,1,6)\end{array}\] linearly independent in \(R^{4}\)? **3.** Find a basis for the subspace of \(R^{4}\) spanned by the four vectors of Exercise 2. **4.** Show that the vectors \[\alpha_{1}=(1,0,-1),\quad\ \ \alpha_{2}=(1,2,1),\quad\ \ \alpha_{3}=(0,-3,2)\] form a basis for \(R^{3}\). Express each of the standard basis vectors as linear combinations of \(\alpha_{1}\), \(\alpha_{2}\) and \(\alpha_{3}\). **5.** Find three vectors in \(R^{2}\) which are linearly dependent, and are such that any two of them are linearly independent. **6.** Let \(V\) be the vector space of all \(2\times 2\) matrices over the field \(F\). Prove that \(V\) has dimension \(4\) by exhibiting a basis for \(V\) which has four elements. **7.** Let \(V\) be the vector space of Exercise 6. Let \(W_{1}\) be the set of matrices of the form \[\begin{bmatrix}x&-x\\ y&z\end{bmatrix}\] and let \(W_{2}\) be the set of matrices of the form \[\begin{bmatrix}a&b\\ -a&c\end{bmatrix}.\] 