

**6.**: Let \(T\) be the linear operator on \(R^{4}\) which is represented in the standard ordered basis by the matrix

\[\begin{bmatrix}0&0&0&0^{7}\\ a&0&0&0\\ 0&b&0&0\\ 0&0&c&0\end{bmatrix}.\]

Under what conditions on \(a\), \(b\), and \(c\) is \(T\) diagonalizable?

**7.**: Let \(T\) be a linear operator on the \(n\)-dimensional vector space \(V\), and suppose that \(T\) has \(n\)_distinct_ characteristic values. Prove that \(T\) is diagonalizable.

**8.**: Let \(A\) and \(B\) be \(n\times n\) matrices over the field \(F\). Prove that if \((I-AB)\) is invertible, then \(I-BA\) is invertible and

\[(I-BA)^{-1}=I+B(I-AB)^{-1}A.\]

**9.**: Use the result of Exercise 8 to prove that, if \(A\) and \(B\) are \(n\times n\) matrices over the field \(F\), then \(AB\) and \(BA\) have precisely the same characteristic values in \(F\).

**10.**: Suppose that \(A\) is a \(2\times 2\) matrix with real entries which is symmetric (\(A^{t}=A\)). Prove that \(A\) is similar over \(R\) to a diagonal matrix.

**11.**: Let \(N\) be a \(2\times 2\) complex matrix such that \(N^{2}=0\). Prove that either \(N=0\) or \(N\) is similar over \(C\) to

\[\begin{bmatrix}0&0\\ 1&0\end{bmatrix}.\]

**12.**: Use the result of Exercise 11 to prove the following: If \(A\) is a \(2\times 2\) matrix with complex entries, then \(A\) is similar over \(C\) to a matrix of one of the two types

\[\begin{bmatrix}a&0^{7}\\ 0&b\end{bmatrix}\quad\begin{bmatrix}a&0\\ 1&a\end{bmatrix}.\]

**13.**: Let \(V\) be the vector space of all functions from \(R\) into \(R\) which are continuous, i.e., the space of continuous real-valued functions on the real line. Let \(T\) be the linear operator on \(V\) defined by

\[(T\!f)(x)=\int_{0}^{\pi}f(t)\,dt.\]

Prove that \(T\) has no characteristic values.

**14.**: Let \(A\) be an \(n\times n\)_diagonal_ matrix with characteristic polynomial

\[(x-c_{1})^{d_{1}}\cdots(x-c_{k})^{d_{k}},\]

where \(c_{1}\), \(\cdots\), \(c_{k}\) are distinct. Let \(V\) be the space of \(n\times n\) matrices \(B\) such that \(AB=BA\). Prove that the dimension of \(V\) is \(d_{1}^{2}+\cdots+d_{k}^{2}\).

**15.**: Let \(V\) be the space of \(n\times n\) matrices over \(F\). Let \(A\) be a fixed \(n\times n\) matrix over \(F\). Let \(T\) be the linear operator 'left multiplication by \(A^{\prime}\) on \(V\). Is it true that \(A\) and \(T\) have the same characteristic values?

### Annihilating Polynomials

In attempting to analyze a linear operator \(T\), one of the most useful things to know is the class of polynomials which annihilate \(T\). Specifically,