If \(V\) is a finite-dimensional (real or complex) vector space and if \((\cdot\,|\,\cdot\,)\) is an inner product on \(V\), there is an associated class of positive linear operators on \(V\). Via (9-10) there is a one-one correspondence between that class of positive operators and the collection of all positive forms on \(V\). We shall use the exercises for this section to emphasize the relationships between positive operators, positive forms, and positive matrices. The following summary may be helpful.

If \(A\) is an \(n\,\times\,n\) matrix over the field of complex numbers, the following are equivalent.

(1) \(A\) is positive, i.e., \(\sum\limits_{j}\sum\limits_{k}A_{k}x_{j}\overline{x}_{k}\,>0\) whenever \(x_{1},\ldots,x_{n}\) are complex numbers, not all 0.

(2) \((X|Y)\,=\,Y^{*}A\,X\) is an inner product on the space of \(n\,\times\,1\) complex matrices.

(3) Relative to the standard inner product \((X|Y)\,=\,Y^{*}X\) on \(n\,\times\,1\) matrices, the linear operator \(X\,\xrightarrow{}\,AX\) is positive.

(4) \(A\,=\,P^{*}P\) for some invertible \(n\,\times\,n\) matrix \(P\) over \(C\).

(5) \(A\,=\,A^{*}\), and the principal minors of \(A\) are positive.

If each entry of \(A\) is real, these are equivalent to:

(6) \(A\,=\,A^{\,\prime}\), and \(\sum\limits_{j}\sum\limits_{k}A_{k}x_{j}x_{k}\,>0\) whenever \(x_{1},\ldots,x_{n}\) are real numbers not all 0.

(7) \((X|Y)\,=\,Y^{\,\prime}AX\) is an inner product on the space of \(n\,\times\,1\) real matrices.

(8) Relative to the standard inner product \((X|Y)\,=\,Y^{\,\prime}X\) on \(n\,\times\,1\) real matrices, the linear operator \(X\,\xrightarrow{}\,AX\) is positive.

(9) There is an invertible \(n\,\times\,n\) matrix \(P\), with real entries, such that \(A\,=\,P^{\,\prime}P\).

### Exercises

**1.** Let \(V\) be \(C^{\sharp}\), with the standard inner product. For which vectors \(\alpha\) in \(V\) is there a positive linear operator \(T\) such that \(\alpha\,=\,T\epsilon_{1}\)?

**2.** Let \(V\) be \(R^{\sharp}\), with the standard inner product. If \(\theta\) is a real number, let \(T\) be the linear operator 'rotation through \(\theta\),'

\[T\theta(x_{1},x_{2})\,=\,(x_{1}\cos\theta-x_{2}\sin\theta,x_{1}\sin\theta+x_ {2}\cos\theta).\]

For which values of \(\theta\) is \(T\theta\) a positive operator?

**3.** Let \(V\) be the space of \(n\,\times\,1\) matrices over \(C\), with the inner product \((X|Y)\,=\,Y^{*}GX\) (where \(G\) is an \(n\,\times\,n\) matrix such that this is an inner product). Let \(A\) be an \(n\,\times\,n\) matrix and \(T\) the linear operator \(T(X)\,=\,AX\). Find \(T^{*}\). If \(Y\) is a fixed element of \(V\), find the element \(Z\) of \(V\) which determines the linear functional \(X\,\xrightarrow{}\,Y^{*}X\). In other words, find \(Z\) such that \(Y^{*}X\,=\,(X|Z)\) for all \(X\) in \(V\).

 