(iii) \(Z(\alpha_{i};T)\) _is orthogonal to \(Z(\alpha_{k};T)\) when \(j\neq k\). Furthermore, the integer \(r\) and the annihilators \(e_{1},\ldots,e_{r}\) are uniquely determined by conditions_ (i) _and_ (ii) _and the fact that no \(\alpha_{k}\) is \(0\)._

_Corollary. If \(A\) is a normal matrix with real_ (_complex_) _entries, then there is a real orthogonal_ (_unitary_) _matrix \(P\) such that \(P^{-1}AP\) is in rational canonical form._

It follows that two normal matrices \(A\) and \(B\) are unitarily equivalent if and only if they have the same rational form; \(A\) and \(B\) are orthogonally equivalent if they have real entries and the same rational form.

On the other hand, there is a simpler criterion for the unitary equivalence of normal matrices and normal operators.

_Definitions. Let \(V\) and \(V^{\prime}\) be inner product spaces over the same field. A linear transformation_

\[U\colon V\to V^{\prime}\]

_is called a_ **unitary transformation** _if it maps \(V\) onto \(V^{\prime}\) and preserves inner products. If \(T\) is a linear operator on \(V\) and \(T^{\prime}\) a linear operator on \(V^{\prime}\), then \(T\) is_ **unitarily equivalent** _to \(T^{\prime}\) if there exists a unitary transformation \(U\) of \(V\) onto \(V^{\prime}\) such that_

\[UTU^{-1}=T^{\prime}.\]

_Lemma. Let \(V\) and \(V^{\prime}\) be finite-dimensional inner product spaces over the same field. Suppose \(T\) is a linear operator on \(V\) and that \(T^{\prime}\) is a linear operator on \(V^{\prime}\). Then \(T\) is unitarily equivalent to \(T^{\prime}\) if and only if there is an orthonormal basis \(\otimes\) of \(V\) and an orthonormal basis \(\otimes^{\prime}\) of \(V^{\prime}\) such that_

\[[T]_{\otimes}=[T^{\prime}]_{\otimes^{\prime}}.\]

_Proof._ Suppose there is a unitary transformation \(U\) of \(V\) onto \(V^{\prime}\) such that \(UTU^{-1}=T^{\prime}\). Let \(\otimes=\{\alpha_{1},\ldots,\alpha_{n}\}\) be any (ordered) orthonormal basis for \(V\). Let \(\alpha^{\prime}_{j}=U\alpha_{j}\) (\(1\leq j\leq n\)). Then \(\otimes^{\prime}=\{\alpha^{\prime}_{1},\ldots,\)\(\alpha^{\prime}_{n}\}\) is an orthonormal basis for \(V^{\prime}\) and setting

\[T\alpha_{j}=\sum_{k=1}^{n}A_{kj}\alpha_{k}\]

we see that

\[\eqalign{T^{\prime}\alpha^{\prime}_{j}&=UT\alpha_{j}\cr&=\sum_{k}A_{kj}U\alpha _{k}\cr&=\sum_{k}A_{kj}\alpha^{\prime}_{k}\cr}\]

Hence \([T]_{\otimes}=A=[T^{\prime}]_{\otimes^{\prime}}\).

 