
For which \(X\) does there exist a scalar \(c\) such that \(AX=cX\)?

* Discover whether \[A=\begin{bmatrix}1&2&3&4\\ 0&2&3&4\\ 0&0&3&4\\ 0&0&0&4\end{bmatrix}\] is invertible, and find \(A^{-1}\) if it exists.
* Suppose \(A\) is a \(2\times 1\) matrix and that \(B\) is a \(1\times 2\) matrix. Prove that \(C=AB\) is not invertible.
* Let \(A\) be an \(n\times n\) (square) matrix. Prove the following two statements:
* If \(A\) is invertible and \(AB=0\) for some \(n\times n\) matrix \(B\), then \(B=0\).
* If \(A\) is not invertible, then there exists an \(n\times n\) matrix \(B\) such that \(AB=0\) but \(B\neq 0\).
* Let \[A=\begin{bmatrix}a&b\\ c&d\end{bmatrix}.\] Prove, using elementary row operations, that \(A\) is invertible if and only if \((ad-bc)\neq 0\).
* An \(n\times n\) matrix \(A\) is called **upper-triangular** if \(A_{ii}=0\) for \(i>j\), that is, if every entry below the main diagonal is \(0\). Prove that an upper-triangular (square) matrix is invertible if and only if every entry on its main diagonal is different from \(0\).
* Prove the following generalization of Exercise 6. If \(A\) is an \(m\times n\) matrix, \(B\) is an \(n\times m\) matrix and \(n<m\), then \(AB\) is not invertible.
* Let \(A\) be an \(m\times n\) matrix. Show that by means of a finite number of elementary row and/or column operations one can pass from \(A\) to a matrix \(R\) which is both 'row-reduced eehelon' and 'column-reduced eehelon,' i.e., \(R_{ii}=0\) if \(i\neq j\), \(R_{ii}=1\), \(1\leq i\leq r\), \(R_{ii}=0\) if \(i>r\). Show that \(R=PAQ\), where \(P\) is an invertible \(m\times m\) matrix and \(Q\) is an invertible \(n\times n\) matrix.
* The result of Example 16 suggests that perhaps the matrix \[A=\begin{bmatrix}1&\frac{1}{2}&\cdots&\frac{1}{n}\\ 1&1&\\ \frac{2}{2}&\frac{3}{3}&\cdots&\frac{1}{n+1}\\ \vdots&\vdots&&\vdots\\ \frac{1}{n}&\frac{1}{n+1}&\cdots&\frac{1}{2n-1}\end{bmatrix}\] is invertible and \(A^{-1}\) has integer entries. Can you prove that? 