

**Exercises**

**1.** Give a reasonable definition of a non-negative \(n\times n\) matrix, and then prove that such a matrix has a unique non-negative square root.

**2.** Let \(A\) be an \(n\times n\) matrix with complex entries such that \(A^{*}=-A\), and let \(B=e^{A}\). Show that

1. \(\det B=e^{\mathrm{tr}\,A}\);
2. \(B^{*}=e^{-A}\);
3. \(B\) is unitary.

**3.** If \(U\) and \(T\) are normal operators which commute, prove that \(U+T\) and \(UT\) are normal.

**4.** Let \(T\) be a linear operator on the finite-dimensional complex inner product space \(V\). Prove that the following ten statements about \(T\) are equivalent.

1. \(T\) is normal.
2. \(\|T\alpha\|=\|T^{*}\alpha\|\) for every \(\alpha\) in \(V\).
3. \(T=T_{1}+iT_{2}\), where \(T_{1}\) and \(T_{2}\) are self-adjoint and \(T_{1}T_{2}=T_{2}T_{1}\).
4. If \(\alpha\) is a vector and \(c\) a scalar such that \(T\alpha=c\alpha\), then \(T^{*}\alpha=\varepsilon\alpha\).
5. There is an orthonormal basis for \(V\) consisting of characteristic vectors [for \(T\).
6. There is an orthonormal basis \(\otimes\) such that \([T]_{\otimes}\) is diagonal.
7. There is a polynomial \(g\) with complex coefficients such that \(T^{*}=g(T)\).
8. Every subspace which is invariant under \(T\) is also invariant under \(T^{*}\).
9. \(T=NU\), where \(N\) is non-negative, \(U\) is unitary, and \(N\) commutes with \(U\).
10. \(T=e_{1}E_{1}+\cdots+e_{k}E_{k}\), where \(I=E_{1}+\cdots+E_{k}\), \(E_{i}E_{j}=0\) for \(i\neq j\), and \(E_{i}^{2}=E_{i}=E_{i}^{*}\).

**5.** Use Exercise 3 to show that any commuting family of normal operators (not necessarily diagonalizable ones) on a finite-dimensional inner product space generates a commutative self-adjoint algebra of normal operators.

**6.** Let \(V\) be a finite-dimensional complex inner product space and \(U\) a unitary operator on \(V\) such that \(U\alpha=\alpha\) implies \(\alpha=0\). Let

\[f(z)=i\,\frac{(1+z)}{(1-z)},\qquad z\neq 1\]

and show that

1. \(f(U)=i(I+U)(I-U)^{-1}\);
2. \(f(U)\) is self-adjoint;
3. for every self-adjoint operator \(T\) on \(V\), the operator \[U=(T-iI)(T+iI)^{-1}\]

is unitary and such that \(T=f(U)\).

**7.** Let \(V\) be the space of complex \(n\times n\) matrices equipped with the inner product

\[(A|B)=\mathrm{tr}\ (AB^{*}).\]