Sec. 5.2

1. (adj \(A\))\(A\) = \(A\)(adj \(A\)) = \((\det A)I\);
2. \(\det\) (adj \(A\)) = \(\det\) (\(A\));
3. \(\adj\) (\(A^{\prime}\)) = \((\adj A)^{\prime}\).

(\(A^{\prime}\) denotes the transpose of \(A\).)

1. Let \(A\) be a \(2\times 2\) matrix over a field \(F\). Show that \(A\) is invertible if and only if \(\det A\neq 0\). When \(A\) is invertible, give a formula for \(A^{-1}\).
2. Let \(A\) be a \(2\times 2\) matrix over a field \(F\), and suppose that \(A^{2}=0\). Show for each scalar \(c\) that \(\det\) (\(cI-A\)) = \(c^{2}\).
3. Let \(K\) be a subfield of the complex numbers and \(n\) a positive integer. Let \(j_{1}\), ..., \(j_{n}\) and \(k_{1}\), ..., \(k_{n}\) be positive integers not exceeding \(n\). For an \(n\times n\) matrix \(A\) over \(K\) define \[D(A)=A(j_{1},k_{1})A(j_{2},k_{3})\ \cdots\ A(j_{n},k_{n}).\]

Prove that \(D\) is \(n\)-linear if and only if the integers \(j_{1}\), ..., \(j_{n}\) are distinct.
4. Let \(K\) be a commutative ring with identity. Show that the determinant function on \(2\times 2\) matrices \(A\) over \(K\) is alternating and \(2\)-linear as a function of the columns of \(A\).
5. Let \(K\) be a commutative ring with identity. Define a function \(D\) on \(3\times 3\) matrices over \(K\) by the rule \[D(A)=A_{11}\det\begin{bmatrix}A_{21}&A_{22}\\ A_{22}&A_{22}\end{bmatrix}-A_{12}\det\begin{bmatrix}A_{11}&A_{22}\\ A_{11}&A_{22}\end{bmatrix}+A_{11}\det\begin{bmatrix}A_{12}&A_{22}\\ A_{22}&A_{22}\end{bmatrix}.\] Show that \(D\) is alternating and \(3\)-linear as a function of the columns of \(A\).
6. Let \(K\) be a commutative ring with identity. Show that the determinant function on \(2\times 2\) matrices \(A\) over \(K\) is alternating and \(2\)-linear as a function of the columns of \(A\).
7. Let \(K\) be a commutative ring with identity. Show that the determinant function on \(2\times 2\) matrices \(A\) over \(K\) is alternating and \(2\)-linear as a function of the columns of \(A\).
8. Let \(K\) be a commutative ring with identity. Define a function \(D\) 