

**Theorem 22**.: _Let \(\mathrm{V}\) be a finite-dimensional complex inner product space and \(\mathrm{T}\) a normal operator on \(\mathrm{V}\). Then \(\mathrm{V}\) has an orthonormal basis consisting of characteristic vectors for \(\mathrm{T}\)._

Again there is a matrix interpretation.

**Corollary**.: _For every normal matrix \(\Lambda\) there is a unitary matrix \(\mathrm{P}\) such that \(\mathrm{P}^{-1}\mathrm{AP}\) is a diagonal matrix._

### Exercises

**1.** For each of the following real symmetric matrices \(A\), find a real orthogonal matrix \(P\) such that \(P^{t}AP\) is diagonal.

\[\begin{bmatrix}1&1\\ 1&1\end{bmatrix},\quad\begin{bmatrix}1&2\\ 2&1\end{bmatrix},\quad\begin{bmatrix}\cos\theta&\sin\theta\\ \sin\theta&-\cos\theta\end{bmatrix}\]

**2.** Is a complex symmetric matrix self-adjoint? Is it normal?

**3.** For

\[A=\begin{bmatrix}1&2&3\\ 2&3&4\\ 3&4&5\end{bmatrix}\]

there is a real orthogonal matrix \(P\) such that \(P^{t}AP=D\) is diagonal. Find such a diagonal matrix \(D\).

**4.** Let \(V\) be \(C^{2}\), with the standard inner product. Let \(T\) be the linear operator on \(V\) which is represented in the standard ordered basis by the matrix

\[A=\begin{bmatrix}1&i\\ i&1\end{bmatrix}.\]

Show that \(T\) is normal, and find an orthonormal basis for \(V\), consisting of characteristic vectors for \(T\).

**5.** Give an example of a \(2\times 2\) matrix \(A\) such that \(A^{2}\) is normal, but \(A\) is not normal.

**6.** Let \(T\) be a normal operator on a finite-dimensional complex inner product space. Prove that \(T\) is self-adjoint, positive, or unitary according as every characteristic value of \(T\) is real, positive, or of absolute value \(1\). (Use Theorem 22 to reduce to a similar question about diagonal matrices.)

**7.** Let \(T\) be a linear operator on the finite-dimensional inner product space \(V\), and suppose \(T\) is both positive and unitary. Prove \(T=I\).

**8.** Prove \(T\) is normal if and only if \(T=T_{1}+iT_{2}\) where \(T_{1}\) and \(T_{2}\) are self-adjoint operators which commute.

**9.** Prove that a real symmetric matrix has a real symmetric cube root; i.e., if \(A\) is real symmetric, there is a real symmetric \(B\) such that \(B^{2}=A\).

**10.** Prove that every positive matrix is the square of a positive matrix.

