Sec. 9.3 _Positive Forms_ _381_

**4.**: Let \(V\) be a finite-dimensional inner product space. If \(T\) and \(U\) are positive linear operators on \(V\), prove that \((T+U)\) is positive. Give an example which shows that \(TU\) need not be positive.
**5.**: Let

\[A=\begin{bmatrix}1&\frac{1}{2}\\ \frac{1}{2}&\frac{1}{2}\end{bmatrix}.\]

(a) Show that \(A\) is positive.

(b) Let \(V\) be the space of \(2\times 1\) real matrices, with the inner product \((X|Y)=Y^{\prime}{}_{4}X\). Find an orthonormal basis for \(V\), by applying the Gram-Schmidt process to the basis \(\{X_{1},X_{2^{i}}\}\) defined by

\[X_{1}=\begin{bmatrix}1\\ 0\end{bmatrix},\qquad X_{1}=\begin{bmatrix}0\\ 1\end{bmatrix}.\]

(c) Find an invertible \(2\times 2\) real matrix \(P\) such that \(A\approx\boldsymbol{P}^{i}P\).
**6.**: Which of the following matrices are positive?

\[\begin{bmatrix}1&2^{2}\\ 3&4\end{bmatrix},\qquad\begin{bmatrix}1&1&+i\\ 1-i&3\end{bmatrix},\qquad\begin{bmatrix}1&-1&1\\ 2&-1&1\\ 3&-1&1\end{bmatrix},\qquad\begin{bmatrix}1&\frac{1}{2}&\frac{1}{2}\\ \frac{1}{2}&\frac{1}{2}&\frac{1}{2}\end{bmatrix}.\]
**7.**: Give an example of an \(n\times n\) matrix which has all its principal minors positive, but which is not a positive matrix.
**8.**: Does \(((x_{1},x_{2})|(y_{1},y_{2}))=x_{2}\bar{y}_{1}+2x_{2}\bar{y}_{1}+2x_{1}\bar{y} _{1}+x_{2}\bar{y}_{2}\) define an inner product on \(C\)?
**9.**: Prove that every entry on the main diagonal of a positive matrix is positive.
**10.**: Let \(V\) be a finite-dimensional inner product space. If \(T\) and \(U\) are linear operators on \(V\), we write \(T<U\) if \(U-T\) is a positive operator. Prove the following:

1. \(T<U\) and \(U<T\) is impossible.
2. If \(T<U\) and \(U<S\), then \(T<S\).
3. If \(T<U\) and \(0<S\), it need not be that \(ST<SU\).
**11.**: Let \(V\) be a finite-dimensional inner product space and \(E\) the orthogonal projection of \(V\) onto some sub-pace.

1. Prove that, for any positive number \(c\), the operator \(cI+E\) is positive.
2. Express in terms of \(E\) a self-adjoint linear operator \(T\) such that \(T^{*}=I+E\).
**12.**: Let \(n\) be a positive integer and \(A\) the \(n\times n\) matrix

\[A=\begin{bmatrix}1&\frac{1}{2}&\frac{1}{3}&\cdots&\frac{1}{n}\\ \frac{1}{2}&\frac{1}{3}&\frac{1}{4}&\cdots&\frac{1}{n+1}\\ \vdots&\vdots&\vdots&&\vdots\\ \frac{1}{n}&\frac{1}{n+1}&\frac{1}{n+2}&\cdots&\frac{1}{2n-1}\\ \hline\end{bmatrix}.\]

Prove that \(A\) is positive.

 