
Find the characteristic polynomial for \(T\). Consider the cases \(a=b=1;a=b=0\); \(\bullet=0\), \(b=1\). In each of these cases, find the minimal polynomial for \(T\) and non-zero vectors \(\alpha_{1}\), ..., \(\alpha\), which satisfy the conditions of Theorem 3.

**11.** Prove that if \(A\) and \(B\) are \(3\times 3\) matrices over the field \(F\), a necessary and sufficient condition that \(A\) and \(B\) be similar over \(F\) is that they have the same characteristic polynomial and the same minimal polynomial. Give an example which shows that this is false for \(4\times 4\) matrices.

**12.** Let \(F\) be a subfield of the field of complex numbers, and let \(A\) and \(B\) be \(n\times n\) matrices over \(F\). Prove that if \(A\) and \(B\) are similar over the field of complex numbers, then they are similar over \(F\). (_Hint:_ Prove that the rational form of \(A\) is the same whether \(A\) is viewed as a matrix over \(F\) or a matrix over \(C\); likewise for \(B\).)

**13.** Let \(A\) be an \(n\times n\) matrix with complex entries. Prove that if every characteristic value of \(A\) is real, then \(A\) is similar to a matrix with real entries.

**14.** Let \(T\) be a linear operator on the finite-dimensional space \(V\). Prove that there exists a vector \(\alpha\) in \(V\) with this property. If \(f\) is a polynomial and \(f(T)\alpha=0\), then \(f(T)=0\). (Such a vector \(\alpha\) is called a **separating vector** for the algebra of polynomials in \(T\).) When \(T\) has a cyclic vector, give a direct proof that any cyclic vector is a separating vector for the algebra of polynomials in \(T\).

**15.** Let \(F\) be a subfield of the field of complex numbers, and let \(A\) be an \(n\times n\) matrix over \(F\). Let \(p\) be the minimal polynomial for \(A\). If we regard \(A\) as a matrix over \(C\), then \(A\) has a minimal polynomial \(f\) as an \(n\times n\) matrix over \(C\). Use a theorem on linear equations to prove \(\boldsymbol{p}=f\). Can you also see how this follows from the cyclic decomposition theorem?

**16.** Let \(A\) be an \(n\times n\) matrix with _real_ entries such that \(A^{2}+I=0\). Prove that \(n\) is even, and if \(n=2k\), then \(A\) is similar over the field of real numbers to a matrix of the block form

\[\begin{bmatrix}0&-I\\ I&0\end{bmatrix}\]

where \(I\) is the \(k\times k\) identity matrix.

**17.** Let \(T\) be a linear operator on a finite-dimensional vector space \(V\). Suppose that

1. the minimal polynomial for \(T\) is a power of an irreducible polynomial;
2. the minimal polynomial is equal to the characteristic polynomial.

Show that no non-trivial \(T\)-invariant subspace has a complementary \(T\)-invariant subspace.

**18.** If \(T\) is a diagonalizable linear operator, then every \(T\)-invariant subspace has a complementary \(T\)-invariant subspace.

**19.** Let \(T\) be a linear operator on the finite-dimensional space \(V\). Prove that \(T\) has a cyclic vector if and only if the following is true: Every linear operator \(U\) which commutes with \(T\) is a polynomial in \(T\).

**20.** Let \(V\) be a finite-dimensional vector space over the field \(F\), and let \(T\) be a linear operator on \(V\). We ask when it is true that every non-zero vector in \(V\) is a cyclic vector for \(T\). Prove that this is the case if and only if the characteristic polynomial for \(T\) is irreducible over \(F\).

 