See 9.6

\[T=\sum_{j=2}^{k}c_{i}P_{j}\]

where \(c_{2}\), ..., \(c_{k}\) are arbitrary scalars.

(d) \(\alpha=\alpha_{0}\) if and only if for each root \(r\) of \(\alpha\) there exists an operator \(T\) in \(G_{0}\) such that \(r(T)\neq 0\).

### Further Properties of Normal Operators

In Section 8.5 we developed the basic properties of self-adjoint and normal operators, using the simplest and most direct methods possible. In Section 9.5 we considered various aspects of spectral theory. Here we prove some results of a more technical nature which are mainly about normal operators on real spaces.

We shall begin by proving a sharper version of the primary decomposition theorem of Chapter 6 for normal operators. It applies to both the real and complex cases.

**Theorem 17**: _Let \(T\) be a normal operator on a finite-dimensional inner product space \(V\). Let \(p\) be the minimal polynomial for \(T\) and \(p_{i}\), \(\cdots\), \(p_{k}\) its distinct monic prime factors. Then each \(p_{i}\) occurs with multiplicity \(1\) in the factorization of \(p\) and has degree \(1\) or \(2\). Suppose \(W_{i}\) is the null space of \(p_{i}(T)\). Then_

* \(W_{i}\) _is orthogonal to_ \(W_{i}\) _when_ \(i\neq j\)_;_
* \(V=W_{1}\bigoplus\,\cdots\,\bigoplus\,W_{k}\)_;_
* \(W_{i}\) _is invariant under_ \(T\)_, and_ \(p_{i}\) _is the minimal polynomial for the restriction of_ \(T\) _to_ \(W_{i}\)_;_
* _for every_ \(j\)_, there is a polynomial_ \(e_{i}\) _with coefficients in the scalar field such that_ \(e_{i}(T)\) _is the orthogonal projection of_ \(V\) _on_ \(W_{j}\)_._

In the proof we use certain basic facts which we state as lemmas.

**Lemma 1**: _Let \(N\) be a normal operator on an inner product space \(W\). Then the null space of \(N\) is the orthogonal complement of its range._

Suppose \((\alpha|N\beta)=0\) for all \(\beta\) in \(W\). Then \((N^{*}\alpha|\beta)=0\) for all \(\beta\); hence \(N^{*}\alpha=0\). By Theorem 19 of Chapter 8, this implies \(N\alpha=0\). Conversely, if \(N\alpha=0\), then \(N^{*}\alpha=0\), and

\[(N^{*}\alpha|\beta)\,=\,(\alpha|N\beta)\,=\,0\]

for all \(\beta\) in \(W\).

 